---
title: "__Candida Klebsiella RNAseq Analysis Revised__"
author: "Celine Prakash"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: kate
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, out.height = "\\textheight",  out.width = "\\textwidth*1.8",dev = c('png', 'pdf'),fig.path='figures/')
```

```{r setup environment, include=FALSE}
SALMON_DIR="../results/expression/salmon_quant"
METADATA_DIR="../metadata"
ANNOTATION_DIR="../reference"
EXPERIMENT_NAME="candida_klebsiella_rnaseq_analysis_results_without_CaKp3_CaKp4_revised"
OUTPUT_DIR=EXPERIMENT_NAME
dir.create(file.path(OUTPUT_DIR), showWarnings = FALSE)
TFs_list_file="CaOrtholog_23Sep23_Raju_PCT_TFs.txt"
exclude_samples=c("CaKp3", "CaKp4")

# renv::init()
#renv::restore()
#renv::snapshot()
```


## __Project Overview__

The aim of this work is to unveil pathways and biological processes affected by the interaction of C. albicans and K. pneumoniae by comparing the mixed cultures vs single cultures. 


Design:

__Cultures:__ Candida albicans (SC5314) only, Klebsiella pneumoniae ATCC 700603 only or C. albicans and K. pneumoniae

__2 types of media:__ RPMI or M9

__Replicates:__ Four or more replicates of each culture-medium combination


```{r installing specific libraries, include=FALSE}
#installing libraries if not present
## packages_needed<-c("rlang","import","readxl","readr","janitor","here","openxlsx")
#packages_needed<-c("memoise","mime")
#biocond_packages_needed<-c("biocc::BiocFileCache","bioc::tximport","bioc::edgeR","bioc::AnnotationDbi","bioc::biomaRt","bioc::GenomicFeatures","biocc::clusterProfiler")
## plotting_packages<-c("ggfortify","pheatmap","RColorBrewer","ggrepel","eulerr","ggpubr","msa")
# 
# install.packages(packages_needed)
# BiocManager::install("BiocFileCache")
# 
# if (!require("BiocManager", quietly = TRUE))
#       install.packages("BiocManager")
#    BiocManager::install(version = '3.16')
# install.packages(biocond_packages_needed)
#   BiocManager::install("topGO")
# 
#  BiocManager::install("remotes")
#  install.packages("devtools")
#  devtools::install_github("montilab/hypeR")

 
# install.packages(plotting_packages)

#renv::record("renv@1.0.5")
# #finish by taking snapshot of libraries
# renv::snapshot()
#renv::activate()
#renv::restore(packages = "renv")

#devtools::install_github("YuLab-SMU/ggtree")
#BiocManager::install("clusterProfiler")
#BiocManager::install("pathview")
#BiocManager::install("ComplexHeatmap")
# BiocManager::install("car")
```

```{r loading libraries, include=FALSE}

#importing key functions
import::from(magrittr, "%>%")
import::from(.from=readxl, read_excel)
import::from(.from=janitor, make_clean_names, remove_empty)
import::from(.from=stringr, str_detect)

library(markdown)
library(tibble)
library(ggplot2)
library(gridExtra)
library(openxlsx)
library(biomaRt)
library(dplyr)
library(ggrepel)
library(topGO)
library(KEGGREST)
library(clusterProfiler)
library(pathview)
library(enrichplot)
library(ComplexHeatmap)
library(RColorBrewer)
library(circlize)
library(plyr)
library(ggfortify)
library(cowplot)
library(car)

```

```{r downldtables, include=FALSE}

#downloadable tables
library(DT)
create_dt <- function(x){
 DT::datatable(x,
  extensions = 'Buttons',
  options = list(dom = 'Blfrtip',
                 buttons = c('copy', 'csv', 'excel'),
                 lengthMenu = list(c(10,25,50,-1),
                                   c(10,25,50,"All"))))
}
```

```{r functions,echo=FALSE}

save_xlsx<-function(x){
  write.xlsx(x,file=file.path(OUTPUT_DIR,paste(deparse(substitute(x)),".xlsx",sep="")),quote=FALSE,rowNames=TRUE)
}

##Venn diagram gene lists functions taken from https://stackoverflow.com/questions/23559371/get-the-list-of-items-in-venn-diagram

Intersect <- function (x) {  
  # Multiple set version of intersect
  # x is a list
  if (length(x) == 1) {
    unlist(x)
  } else if (length(x) == 2) {
    intersect(x[[1]], x[[2]])
  } else if (length(x) > 2){
    intersect(x[[1]], Intersect(x[-1]))
  }
}

Union <- function (x) {  
  # Multiple set version of union
  # x is a list
  if (length(x) == 1) {
    unlist(x)
  } else if (length(x) == 2) {
    union(x[[1]], x[[2]])
  } else if (length(x) > 2) {
    union(x[[1]], Union(x[-1]))
  }
}

Setdiff <- function (x, y) {
  # Remove the union of the y's from the common x's. 
  # x and y are lists of characters.
  xx <- Intersect(x)
  yy <- Union(y)
  setdiff(xx, yy)
}

write_venn_genelists <- function(signatures,sig_names_interest,outbasedir){
  # Create a list of all the combinations 
# Taken from https://stackoverflow.com/questions/23559371/get-the-list-of-items-in-venn-diagram
combs <- 
  unlist(lapply(1:length(sig_names_interest), 
                function(j) combn(sig_names_interest, j, simplify = FALSE)),
         recursive = FALSE)
names(combs) <- sapply(combs, function(i) paste0(i, collapse = "__"))
elements <- 
  lapply(combs, function(i) Setdiff(signatures[sig_names_interest][i], signatures[sig_names_interest][setdiff(names(signatures[sig_names_interest]), i)]))
d<-file.path(outbasedir)
if(!dir.exists(d)){
  dir.create(d)
}
of<-lapply(names(elements),function(e){
  filename=paste(e,".txt",sep="")
  write.table(elements[[e]], file=file.path(outbasedir,filename),quote=FALSE,sep="\t",row.names = F,col.names = F)
})
}

#modified from Peter R's plotDegResults Volcano function
plotVolcano <- function(results_obj=NULL, genes_of_interest=NULL, plot_title=NULL,FC_thresh=log2(2),adj_p_value_thresh=0.05){
  forPlot<-results_obj
  forPlot$signif_DE <- "NO"
  forPlot$signif_DE[forPlot$logFC>FC_thresh&forPlot$adj.P.Val<adj_p_value_thresh] <- "UP"
  forPlot$signif_DE[forPlot$logFC<(-FC_thresh)&forPlot$adj.P.Val<adj_p_value_thresh] <- "DOWN"
  forPlot$signif_DE <- factor(forPlot$signif_DE, levels = c("NO", "DOWN", "UP"))
  table(forPlot$signif_DE)
  signif_volcanoPlot <- ggplot(data = forPlot, aes(x = logFC, y = -log10(adj.P.Val), col=signif_DE)) +
    geom_point() +
    ggrepel::geom_text_repel(data=forPlot[which(forPlot$symbol %in% genes_of_interest),], aes(label = forPlot$symbol[which(forPlot$symbol %in% genes_of_interest)]),show.legend = FALSE,max.overlaps=20, label.size = 0.05,force=20,force_pull=0.2,direction="both") +
    scale_color_manual(values=c(DOWN=brewer.pal(n = 8, name = "RdBu")[8], UP=brewer.pal(n = 8, name = "RdBu")[1], NO = "grey")) +
    theme_bw(base_size = 14) +
    theme(panel.border = element_blank()) +
    theme(axis.line = element_line(color = 'black')) +
    labs(x = "log2FC") +
    geom_hline(yintercept=-log10(adj_p_value_thresh),linetype="dashed") +
    geom_vline(xintercept=c(-FC_thresh,FC_thresh),linetype="dashed") +
    ggtitle(plot_title)
}

# plotVolcano2 
plotVolcano2 <- function(results_obj=NULL, genes_of_interest=NULL, plot_title=NULL,
                        FC_thresh=log2(2), adj_p_value_thresh=0.05) {
  forPlot <- results_obj
  forPlot$signif_DE <- "NO"
  forPlot$signif_DE[forPlot$logFC >  FC_thresh & forPlot$adj.P.Val < adj_p_value_thresh] <- "UP"
  forPlot$signif_DE[forPlot$logFC < -FC_thresh & forPlot$adj.P.Val < adj_p_value_thresh] <- "DOWN"
  forPlot$signif_DE <- factor(forPlot$signif_DE, levels = c("NO", "DOWN", "UP"))

  label_data          <- forPlot[forPlot$symbol %in% genes_of_interest, ]
  label_data$nudge_x  <- ifelse(label_data$logFC > 0,  0.8, -0.8)
  label_data$nudge_y  <- 0.5

  ggplot(data = forPlot, aes(x = logFC, y = -log10(adj.P.Val), col = signif_DE)) +
    geom_point() +
    ggrepel::geom_text_repel(
      data               = label_data,
      aes(label          = symbol),
      nudge_x            = label_data$nudge_x,
      nudge_y            = label_data$nudge_y,
      show.legend        = FALSE,
      max.overlaps       = 20,
      force              = 20,
      force_pull         = 0.2,
      direction          = "both",
      point.padding      = 0.4,
      min.segment.length = 0,
      bg.color           = "white",
      bg.r               = 0.1,
      size               = 3.5
    ) +
    scale_color_manual(values = c(
      DOWN = brewer.pal(n = 8, name = "RdBu")[8],
      UP   = brewer.pal(n = 8, name = "RdBu")[1],
      NO   = "grey"
    )) +
    theme_bw(base_size = 14) +
    theme(panel.border = element_blank(),
          axis.line    = element_line(color = "black")) +
    labs(x = "log2FC") +
    geom_hline(yintercept = -log10(adj_p_value_thresh), linetype = "dashed") +
    geom_vline(xintercept = c(-FC_thresh, FC_thresh),   linetype = "dashed") +
    ggtitle(plot_title)
}

#Colour function
col_fun <- colorRamp2(
  breaks = seq(-3, 3, length.out = 8),
  colors = rev(brewer.pal(n = 8, name = "RdBu"))
)

#Gene heatmap
Gene_Heatmap<-function(matlog2fc,matadjpval,title){
  H<-Heatmap(
  matlog2fc,
  cell_fun = function(j, i, x, y, w, h, fill) {
    if (matadjpval[i,j] < 0.05) {
      grid.text("*", x, y)
    }
  },
    cluster_columns = FALSE,
    col =  col_fun,
    name = 'Log2FC',
    column_title = title,
    border = "white",
    rect_gp = gpar(col = "white"),
    width = unit(5, "mm"),
    height = unit(nrow(matlog2fc) * 5, "mm"),
    heatmap_legend_param = list(title = "Log2FC", at = seq(-3, 3, 1))
  )
  return(H)
}
```

```{r function to get uniquely significant,echo=FALSE}
get_uniquely_significant<-function(mhyp,indices1,indices2){
  significant_list<-lapply(names(mhyp$data),function(l){mhyp$data[[l]]$data$label[which(mhyp$data[[l]]$data$fdr<=0.05)]})
  names(significant_list)<-names(mhyp$data)
  comparisons1<-names(significant_list)[indices1]
  subset_significant_list1<-unique(unlist(significant_list[comparisons1]))
  comparisons2<-names(significant_list)[indices2]
  subset_significant_list2<-unique(unlist(significant_list[comparisons2]))
  unique_to_1<-subset_significant_list1[!subset_significant_list1 %in% subset_significant_list2]
  unique_to_2<-subset_significant_list2[!subset_significant_list2 %in% subset_significant_list1]
  return(list(unique_to_1,unique_to_2))
}
get_common_significant<-function(mhyp,index1,index2){
  significant_list<-lapply(names(mhyp$data),function(l){mhyp$data[[l]]$data$label[which(mhyp$data[[l]]$data$fdr<=0.05)]})
  names(significant_list)<-names(mhyp$data)
  comparisons1<-names(significant_list)[index1]
  subset_significant_list1<-unique(unlist(significant_list[comparisons1]))
  comparisons2<-names(significant_list)[index2]
  subset_significant_list2<-unique(unlist(significant_list[comparisons2]))
  common<-subset_significant_list1[subset_significant_list1 %in% subset_significant_list2]
  comparisons3<-names(significant_list)[which(!seq(length(mhyp$data))%in%c(index1,index2))]
  subset_significant_list3<-unique(unlist(significant_list[comparisons3]))
  common_unique<-common[!common %in% subset_significant_list3]
  return(list(common,common_unique))
}
```


```{r TopGO related functions, echo=FALSE}


##Function to filter topGO results by GO id
subset_by_go<-function(results,go_ids){
  return(results[go_ids,])
}

##Sort topGO results by increasing (significant) adj p-value

#  calculate the pvalue average, on condition that it is significant
conditional_average<-function(x,upper_threshold){
  mean(x[x<upper_threshold])
}

# sorting function
sort_sig_adjpval<-function(filtered_results_list,upper_threshold=upper_threshold){
  
  #calculate the average p-values (for significant)
  p_values<-as.data.frame(lapply(filtered_results_list,function(l){l['adj.Pvalue']}))
  go_adjp_ave<-apply(p_values,1,conditional_average,upper_threshold=upper_threshold)

  # sort result by adj p-val average
  signif_result_list<-lapply(filtered_results_list,function(l){
  return(l[names(sort(go_adjp_ave)),])
  })
  names(signif_result_list)<-names(filtered_results_list)
  return(signif_result_list)
}

##reverse log transform function from HypeR
reverselog_trans <- function(base=exp(1)) {
    trans <- function(x) -log(x, base)
    inv <- function(x) base^(-x)
    scales::trans_new(paste0("reverselog-", format(base)), trans, inv, 
              scales::log_breaks(base=base), 
              domain=c(1e-100, Inf))
}


##Plot function for multiple comparisons (that looks like HypeR output)
dotplot_comparisons<-function(signif_result_list,FDR_thres){
  # prepare dataframe with the ratio of Significant/Annotated and the FDR values from each signature
  Comparison<-c()
  Term<-c()
  Ratio<-c()
  FDR<-c()
  for(i in seq(length(signif_result_list))){
    data<-signif_result_list[[i]]
    Comparison<-c(Comparison,rep(names(signif_result_list)[i],length(data$Term)))
    Term<-c(Term,data$Term)
    Ratio<-c(Ratio,data$Significant/data$Annotated)
    FDR<-c(FDR,data$adj.Pvalue)
  }
  Term <- factor(Term, levels=rev(unique(Term)))
  sigdataplot<-data.frame(Comparison=Comparison,Term=Term,Ratio=Ratio,FDR=FDR)
  sigdataplot_fil<-sigdataplot[sigdataplot$FDR<FDR_thres,]
  sigplot<-ggplot(sigdataplot_fil, aes(x = Comparison, y = Term, size = Ratio, color = FDR)) +
  geom_point() +
  scale_color_gradient2(
    low = brewer.pal(8, "RdBu")[2],
    mid = "white",
    high = brewer.pal(8, "RdBu")[7],
    midpoint = 0.055,
    limits = c(0,0.07)
  ) +
  scale_x_discrete(limits = names(GOdata_list)) +
  theme_linedraw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
  return(sigplot)
}

##Function to add gene names to GO data
genes_per_pathway<-function(GOdata_list,signatures,topGO_results){
  ##get annotated genes for each GO term
  ann.genes=genesInTerm(GOdata_list[[1]]) #GO terms are the same for all GOdata. identical(ann.genes,ann.genes_2)
  
  ##get only defined differentially expressed genes in allGO
  tableGO.genes_list<-lapply(names(GOdata_list),function(comp){
    ANOTATION = lapply(ann.genes,function(x) x[x %in% signatures[[comp]]])
    
  ##comma separated merged gene list
    genes_per_go<-as.data.frame(sapply(ANOTATION, paste, collapse = ','))
  names(genes_per_go)<-"Gene_Symbols"
  return(genes_per_go)
  })
  
  ##add this to topgo results
  for (i in seq(length(topGO_results$result_list))){
      topGO_results$result_list[[i]]$Gene_Symbols<-tableGO.genes_list[[i]][topGO_results$result_list[[i]]$GO.ID,1]
  }
  return(topGO_results)
}
  


```

```{r TopGO wrapper,echo=FALSE}

##Wrapper starts here
topGO_wrapper<-function(GOdata_list,test.stat,size_thres=1000,Top_pathways=50,FDR_thres=0.05){

  result_list<-lapply(GOdata_list,function(GOdata){
  result <- getSigGroups(GOdata, test.stat)
  allResWt01<-GenTable(GOdata, Pvalue = result,topNodes=length(usedGO(GOdata)),numChar=10000)

  #3. filter out pathways that are very large (general)
  if (!is.na(size_thres)){
  allResWt01<-allResWt01[allResWt01$Annotated<size_thres,]
  }
  
  #4. multiple testing correction
  allResWt01$adj.Pvalue<-p.adjust(allResWt01$Pvalue,method = "BH")
  rownames(allResWt01)<-allResWt01$GO.ID
  return(allResWt01)
  })

  #5. get unique list of adjusted p-value significant GO terms
  signif_terms_list<-lapply(result_list,function(l){
  signif_terms<-l[as.numeric(l$adj.Pvalue)<0.05,'GO.ID']
  return(signif_terms)
  })
  all_signif_terms<-unique(unlist(signif_terms_list))
  #function to identify which comparisons have significant adj p-values
  track_signif_comparison<-lapply(signif_terms_list,function(l){
    all_signif_terms%in%l
  })
  #get significant comparisons across for all significant go terms
  comp_signif<-apply(as.data.frame(track_signif_comparison),1,function(x){paste(names(which(x)),collapse = ";")})
  go_ids<-all_signif_terms
  all_signif_terms<-cbind(go_ids,comp_signif)
  
  #6. filter go results by go id
  filtered_results_list<-lapply(result_list,subset_by_go,all_signif_terms[,'go_ids'])
  signif_result_list<-filtered_results_list
  names(signif_result_list)<-names(filtered_results_list)

  #7. plot pathways with FDR values across all signatures
  #sigplot<-dotplot_comparisons(signif_result_list,Top_pathways,FDR_thres)
  sigplot<-dotplot_comparisons(signif_result_list,FDR_thres)
    
  topGO_results<-list(result_list=result_list,all_signif_terms=all_signif_terms,sigplot=sigplot)
  return(topGO_results)
}

## Function to filter topgo results by significant comparisons
filter_topgo_results<-function(comparisons_to_keep,topGO_results){
  selected_go_ids<-c()
  for (comparison in comparisons_to_keep){
    selected_go_ids<-c(selected_go_ids,topGO_results$all_signif_terms[which(topGO_results$all_signif_terms[,'comp_signif']==comparison),'go_ids'])
  }

filtered_topgo_results<-lapply(topGO_results$result_list,subset_by_go,selected_go_ids)
}
```

```{r gsea plot function, echo=FALSE}
geneSetID=1:10
colourlist<-c("#1f77b4","#ff7f0f","#2ba02b","#d62727","#9467bd","#8c564c","#e377c1","#7f7f7f","#bcbd21","#15becf")

#Helper function to remove species from pathway names
strip_species <- function(x) {
  sub(
    " *- *(Candida albicans|Klebsiella pneumoniae subsp\\. pneumoniae MGH 78578 \\(serotype K52\\))$",
    "", x
  )
}

#GSEA plot without the p-value table
modgseaplot<-function(keggGs,geneSetID,colourlist){
  colourlist<-setNames(colourlist,sort(keggGs[geneSetID,"Description"]))
  gsea_plot <-gseaplot2(keggGs, geneSetID,col=colourlist,pvalue_table = F,rel_heights = c(1.75,0.25,0.5),subplots = 1:3)
  gsea_plot[[1]]<-gsea_plot[[1]]+theme(legend.position = "none")
  # p-value table
  pd <- keggGs[geneSetID, c("Description", "pvalue", "p.adjust")]
  pd<-pd[sort(pd$Description,index.return=T)$ix,]
  short_labels <- strip_species(pd$Description)
  dup <- short_labels %in% short_labels[duplicated(short_labels) | duplicated(short_labels, fromLast = TRUE)]
  short_labels[dup] <- pd$Description[dup]
  rownames(pd) <- short_labels
  pd <- pd[,-1]
  for (i in seq_len(ncol(pd))) {
    pd[, i] <- format(pd[, i], digits = 4)
  }
  pvalue_table_grob <- tableGrob(pd)
  colour_mapping <- setNames(colourlist,strip_species(sort(keggGs[geneSetID, "Description"])))
  layout_names <- pvalue_table_grob$layout$name
  rownames_indices <- which(layout_names == "rowhead-fg")  # "rowhead-fg" is often used for row names
  
  # Modify colors of the row names
  for (i in seq_along(rownames_indices)) {
    idx <- rownames_indices[i]
    grob <- pvalue_table_grob$grobs[[idx]]
    rowname <- grob$label
    if (rowname %in% rownames(pd)) {
      # Apply the color
      grob$gp <- gpar(col = colour_mapping[rowname])
      pvalue_table_grob$grobs[[idx]] <- grob  # Update grobs list
    }
  }
    
    final_plot<-plot_grid(print(gsea_plot),pvalue_table_grob,ncol=2,rel_heights= c(0.55,0.45))
  return(final_plot)
}
```

```{r kegg dotplot colour scale, echo=FALSE}
dotplot_scale<-function(keggEnr,title="Kegg Enrichment",font.size=10){
  p<-enrichplot::dotplot(keggEnr,title=title,font.size=font.size)+scale_color_gradient2(
    low = brewer.pal(8, "RdBu")[2],
    mid = "white",
    high = brewer.pal(8, "RdBu")[7],
    midpoint = 0.05,
    limits = c(0,0.07)
)
  return(p)
}
```

## __Metadata__

#### __Table from Tamires with an additional factor made from a combination of group names 1 and 2__

```{r read metadata, include=FALSE}
metadata_raw <- readxl::read_excel(file.path(METADATA_DIR, "table_from_tamires.xlsx"),
                                   col_types = "guess",
                                   na = "NA",
                                   .name_repair = janitor::make_clean_names)

# removing empty rows and columns and samples to be excluded
metadata <- metadata_raw %>%
  janitor::remove_empty(., which = c("rows", "cols"), quiet = FALSE) %>%
  filter(!sample_name_in_report %in% exclude_samples)

```

The following outlier samples have been removed: `r print(exclude_samples)`.

```{r,echo=FALSE}
unique(metadata$group_name_1)
```

```{r,echo=FALSE}
unique(metadata$group_name_2)
```

See additional column "species_media" in table below.

```{r new factor,echo=FALSE}
#Making Klebsiella single samples consistent
metadata$group_name_2[which(metadata$group_name_2=="Klebsiella single")]="Kp single"
#Combination of both factors
metadata$species_media<-paste(metadata$group_name_2,metadata$group_name_1,sep="__")
metadata$species_media<-sub(" ","_",metadata$species_media)
create_dt(metadata)
```


## __Data preprocessing__ 

Data mapping and quantification was done using `salmon quant` with library type specified as "SF". [Publication](https://doi.org/10.1038/nmeth.4197 "Patro, R., Duggal, G., Love, M. et al. Salmon provides fast and bias-aware quantification of transcript expression. Nat Methods 14, 417–419 (2017)."). For information on library type specification see: [library_type](https://salmon.readthedocs.io/en/latest/library_type.html).



### __Import of quantification data__

```{r metadata for salmon import}
sample_name_for_salmon<-sub(" ","",metadata$sample_name_in_report)
salmon_ids <- paste(sample_name_for_salmon,"quant",sep="_")
files <- file.path(SALMON_DIR,salmon_ids,"quant.sf")
file.exists(files)
coldata <- data.frame(files, names = sample_name_for_salmon, stringsAsFactors = FALSE)
names(files)<-sample_name_for_salmon
```

```{r creating tx2gene, eval=TRUE, echo=FALSE, message=FALSE}

#Transcript to genes lookup
tx2gene_file <- file.path(ANNOTATION_DIR, "GCF_000182965.3_ASM18296v3__GCF_000016305.1_ASM1630v1_cds_tx2gene.tsv") # created from cds header lines
tx2gene_gff3 <- readr::read_tsv(tx2gene_file,
                             col_names = c("tx_id", "gene_id"),
                             show_col_types = FALSE)
```

__Importing using tximport__

 `tximport`: [Publication](https://www.nature.com/articles/nmeth.4197 "Charlotte Soneson, Michael I. Love, Mark D. Robinson (2015): Differential analyses for RNA-seq: transcript-level estimates improve gene-level inferences. F1000Research")

```{r tximport}
library(tximport)
txi <- tximport(files, 
                type = "salmon", 
                tx2gene = tx2gene_gff3, 
                txOut = FALSE,
                countsFromAbundance="no",
                ignoreTxVersion=FALSE)
```

## __Candida albicans genes__
```{r subsetting ca}
txi_ca<-list()
txi_ca$abundance<-txi$abundance[grep("CAAL",rownames(txi$abundance)),grep("C",colnames(txi$abundance))]
txi_ca$counts<-txi$counts[grep("CAAL",rownames(txi$counts)),grep("C",colnames(txi$counts))]
txi_ca$length<-txi$length[grep("CAAL",rownames(txi$length)),grep("C",colnames(txi$lengths))]
txi_ca$countsFromAbundance<-txi$countsFromAbundance
```

### __Total reads per library__

```{r ca total reads lib}
barplot_data<-data.frame(
  counts=colSums(txi_ca$counts),
  species=metadata$group_name_2[grep("C",metadata$species_media)],
  media=metadata$group_name_1[grep("C",metadata$species_media)],
  names=metadata$sample_name_in_report[grep("C",metadata$species_media)]
)
barplot_data<-barplot_data[sort(barplot_data$species,decreasing=T,index=T)$ix,]
barplot_data$names<-factor(barplot_data$names,levels=barplot_data$names)
p <- ggplot(data=barplot_data,aes(x=names,y=counts,color=media,fill=species)) + 
      geom_bar(stat="identity") + 
      scale_fill_brewer(palette="Set1") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
      scale_color_brewer(palette="Dark2")
p
```

__Create a DGEList__

```{r ca create DGEList}
suppressPackageStartupMessages(library(edgeR))
y <- DGEList(txi_ca$counts)
```

### __Removing genes that are not expressed__

__Specifying the factors in the experimental design__

```{r ca factors, echo=FALSE}
SPECIES_MEDIA<-metadata$species_media[grep("C",metadata$species_media)]
```

```{r ca design initial}
design<-model.matrix(~0 + SPECIES_MEDIA)
```

__Flter low expressed genes using edgeR's filterByExpr__
```{r ca exressionpfiltering}

keep<-filterByExpr(y,design = design)
filtered_y<-y[keep,]

```

No. of genes before filtering: __`r nrow(y)`__

No. of genes after filtering out low expression genes: __`r nrow(filtered_y)`__

```{r ca plotexpfilter, echo=FALSE}
legend_names=colnames(y)
#plot log2 CPM for genes before and after filtering out low expression genes
#before filtering
lcpm=cpm(y$counts,log = TRUE)
L <- mean(y$samples$lib.size) * 1e-6
M <- median(y$samples$lib.size) * 1e-6
lcpm.cutoff <- log2(10/M + 2/L)
nsamples <- ncol(y)
col <- hcl.colors(nsamples, palette = "zissou1", alpha = NULL, rev = FALSE, fixup = TRUE)
par(mfrow=c(1,2))
plot(density(lcpm[,1]), col=col[1], lwd=2, ylim=c(0,0.26), las=2, main="", xlab="",cex.axis=0.8,cex.lab=0.8)
title(main="A. Raw data", xlab="Log-cpm",cex.main=0.8,cex.lab=0.8)
abline(v=lcpm.cutoff, lty=3)
for (i in 2:nsamples){
  den <- density(lcpm[,i])
  lines(den$x, den$y, col=col[i], lwd=2)
}
legend("topright", legend_names, text.col=col, bty="n",cex=0.4)
#after filtering
lcpm <- cpm(filtered_y, log=TRUE)
plot(density(lcpm[,1]), col=col[1], lwd=2, ylim=c(0,0.26), las=2, main="", xlab="",cex.axis=0.8,cex.lab=0.8)
title(main="B. Filtered data", xlab="Log-cpm",cex.main=0.8,cex.lab=0.8)
abline(v=lcpm.cutoff, lty=3)
for (i in 2:nsamples){
  den <- density(lcpm[,i])
  lines(den$x, den$y, col=col[i], lwd=2)
}
legend("topright", legend_names, text.col=col, bty="n",cex=0.4)
```

### __Expression data__

```{r ca additional gene info}
#genes info lookup
feature_table_file <- file.path(ANNOTATION_DIR, "GCF_000182965.3_ASM18296v3_feature_table.txt") # created from cds header lines
gene_feature_table<- read.table(feature_table_file,
                             header = TRUE,
                             sep="\t",
                             stringsAsFactors = FALSE,
                             comment.char =""
                          )
gene_info<-gene_feature_table[which(gene_feature_table[,1]=="CDS"),c("locus_tag","symbol","name")]
```

Counts per million expression values are included in the table below. 
```{r ca expdata, echo=FALSE}
#create downloadable normalized counts data table
expression_table<-cpm(filtered_y,log=TRUE) #log cpm
dataframe_expression_table_exp<-as.data.frame(expression_table)
rownames(gene_info)<-gene_info$locus_tag
dataframe_expression_table<-merge(gene_info,dataframe_expression_table_exp,by = c(0,0),all.y=TRUE)
rownames(dataframe_expression_table)<-dataframe_expression_table$Row.names
#replace locus_tag columm with row names (as some genes were not in the gene feature table)
dataframe_expression_table$locus_tag<-dataframe_expression_table$Row.names

## change values to non log2
dataframe_expression_table[,grep("^C",colnames(dataframe_expression_table))]<-2^dataframe_expression_table[,grep("^C",colnames(dataframe_expression_table))]
ca_counts_per_million_expression_table<-dataframe_expression_table[,-c(1)]
create_dt(ca_counts_per_million_expression_table)
save_xlsx(ca_counts_per_million_expression_table)
```



## __CA genes: Exploratory data analysis__

### __Top 1000 most variable genes for the exploratory data analysis__

```{r ca cov}
cov_lcpm<-apply(lcpm,1,function(x){sd(x)/mean(x)})
num_highest_cov=1000
keep_2<-names(head(sort(cov_lcpm,decreasing=TRUE),num_highest_cov))
top_lcpm<-lcpm[keep_2,]
ca_top_lcpm<-top_lcpm
```

#### __Heatmap with hierarchical clustering between samples (across all genes)__

`r nrow(top_lcpm)` genes with the highest coefficients of variation were used for the analyses in this section

Euclidean distances between samples were calculated with log2 counts per million (CPM) values across all genes
```{r ca similarityclus, echo=FALSE,fig.height=6,fig.width=9}
library(pheatmap)
library(RColorBrewer)
colors <- colorRampPalette(rev(brewer.pal(9, "Blues")))(255)
fs=9
lcpm_dist<-dist(t(top_lcpm))
lcpm_distMatrix<-as.matrix(lcpm_dist)


heatmapnames <- metadata$sample_name_in_report[grep("C",metadata$species_media)]

# define the annotation
annotation_row = data.frame(
                          culture = factor(metadata$group_name_2[grep("C",metadata$species_media)]),
                          media = factor(metadata$group_name_1[grep("C",metadata$species_media)])
                        )
rownames(annotation_row) = heatmapnames
annotation_colour<-list()
colour_plot<-hcl.colors(5, palette = "zissou1", alpha = NULL, rev = TRUE, fixup = TRUE)
j=1
for (i in c(1:ncol(annotation_row))){
  n=length(levels(annotation_row[,names(annotation_row)[i]]))
  annot_col=colour_plot[j:(j+n-1)]
  j=j+n
  names(annot_col)<-levels(annotation_row[,names(annotation_row)[i]])
  annotation_colour[[i]]<-annot_col  
}
names(annotation_colour)<-names(annotation_row)

rownames(lcpm_distMatrix)<-heatmapnames
colnames(lcpm_distMatrix)<-heatmapnames
par(mar=c(1,1,1,1))
pheatmap(lcpm_distMatrix,
         clustering_distance_rows = lcpm_dist,
         clustering_distance_cols = lcpm_dist,
         col = colors,
         fontsize_row = fs, fontsize_col = fs,
         cellheight = 15, cellwidth = 15,
         annotation_row = annotation_row,
         annotation_colors = annotation_colour,
         angle_col=90
         )
```

#### __Principal Component Analysis (PCA)__

```{r ca PCA generation, message=FALSE, eval=TRUE, include=TRUE, echo=FALSE}
additional_col <- data.frame(culture = factor(metadata$group_name_2[grep("C",metadata$species_media)]),media = factor(metadata$group_name_1[grep("C",metadata$species_media)]), stringsAsFactors =FALSE)
data_for_pca <- cbind(additional_col,t(as.matrix(top_lcpm)))
pca_object <- prcomp(t(as.matrix(top_lcpm)), center = TRUE, scale=TRUE)
summ<-summary(pca_object)
```

About `r round(summ$importance[3,2]*100,1)`% of the total variation is explained in PC1 and PC2 of the PCA.


```{r ca pca plot, echo=FALSE,fig.height=6,fig.width=9}
par(mar=c(3,3,3,3))
p<-autoplot(pca_object,data=data_for_pca,colour='culture',shape='media',label=TRUE,size=3,label.size=4)
p+
  scale_color_manual(values=annotation_colour$culture)
```


## __CA genes: Differential Expression Analysis__

__Comparisons of interest__

The following comparisons are of interest for Candida

	    a) Cakp vs Ca (from RPMI group)
	    b) Cakp vs Kp (from RPMI group)
	    c) Candida_kp vs Candida (from M9 group)
      d) Candida_Kp vs Klebsiella (from M9 group)


__Normalization__

[TMM normalization](https://doi.org/10.1186/gb-2010-11-3-r25 "Robinson, M.D., Oshlack, A. A scaling normalization method for differential expression analysis of RNA-seq data. Genome Biol 11, R25 (2010).") for between sample comparisons  

```{r ca normalization}

# normalize 
dge <- calcNormFactors(filtered_y)
lcpm=cpm(dge$counts,log = TRUE)
boxplot(lcpm, las=2,ylab="log2CountsPerMillion",cex.axis=0.45,ylim=c(-10,20),cex.main=0.8,cex.lab=0.8,pch='.')
```

#### __Heatmap of gene expression for RPMI samples (across all genes)__

```{r ca gene expression heatmap RPMI}
lcpm_heatmap<-lcpm[,which(colnames(lcpm) %in% metadata[which(metadata[,"group_name_1"]=="RPMI"),]$sample_name_in_report)]
lcpm_heatmap<-lcpm_heatmap[apply(lcpm_heatmap,1,sd)>0,] #drop genes invariant across this sample subset (cannot be row-scaled)
p<-pheatmap(lcpm_heatmap, scale = "row",show_rownames = FALSE,heatmap_legend_param = list(title = "Row Z-score"),name="RPMI",col=rev(brewer.pal(n = 8, name = "RdBu")),cellwidth=15)
pdf(file.path(OUTPUT_DIR,"Ca_gene_expression_RPMI_heatmap.pdf"),height=9,width=8)
p
dev.off()
p
```


#### __Heatmap of gene expression for M9 samples (across all genes)__

```{r ca gene expression heatmap M9}
lcpm_heatmap<-lcpm[,which(colnames(lcpm) %in%  gsub(" ", "",metadata[which(metadata[,"group_name_1"]=="M9"),]$sample_name_in_report))]
lcpm_heatmap<-lcpm_heatmap[apply(lcpm_heatmap,1,sd)>0,] #drop genes invariant across this sample subset (cannot be row-scaled)
p<-pheatmap(lcpm_heatmap, scale = "row",show_rownames = FALSE,heatmap_legend_param = list(title = "Row Z-score"),name="M9",col=rev(brewer.pal(n = 8, name = "RdBu")),cellwidth=15)
pdf(file.path(OUTPUT_DIR,"Ca_gene_expression_M9_heatmap.pdf"),height=9,width=8)
p
dev.off()
p
```

__Thresholds__

```{r ca thresholds}
#DEG Thresholds
adj_p_value_thresh=0.05
FC=1.5
FC_thresh=log2(FC)
```

__Transform__

[limma-voom](https://doi.org/10.1093/nar/gkv007 "Ritchie ME, Phipson B, Wu D, Hu Y, Law CW, Shi W, Smyth GK (2015). “limma powers differential expression analyses for RNA-sequencing and microarray studies.” Nucleic Acids Research, 43(7), e47.") transformation to remove heteroscedascity

```{r ca design modif, echo=FALSE}
colnames(design)<-gsub("SPECIES_MEDIA","",colnames(design))
colnames(design)<-gsub(" ","_",colnames(design))
colnames(design)<-gsub("_mix/","",colnames(design))

```

```{r ca voom} 
par(mfrow=c(1,2))
v <- voom(dge, design,plot=T)
fit<-lmFit(v,design)
efit <- eBayes(fit)
plotSA(efit, main="Final model: Mean-variance trend")
```

### __Comparison: A. Cakp vs Ca (from RPMI group)__

#### __Contrast matrix__
```{r ca Contrast cakp vs ca rpmi}
contrast.matrix <- makeContrasts(CaKp_vs_Ca_RPMI = CaKp_mix__RPMI - Ca_single__RPMI, levels=colnames(design))
```

#### __Significant DEGs__
```{r ca DEG cakp vs ca rpmi}
fit_2 <- contrasts.fit(fit,contrast.matrix)
fit_2 <- eBayes(fit_2)

##Benjamini-Hochberg method used to adjust the p-values for multiple testing
results<-topTable(fit_2, adjust="BH",number=nrow(filtered_y))
results_CaKp_vs_Ca_RPMI<-results
#all significant DEGs based on p-value and FC thresholds
results_CaKp_vs_Ca_RPMI_DEGS<-results[which((abs(results$logFC)>FC_thresh&results$adj.P.Val<adj_p_value_thresh)),]
results_CaKp_vs_Ca_RPMI_UPREG<-results[which((results$logFC>FC_thresh&results$adj.P.Val<adj_p_value_thresh)),]
results_CaKp_vs_Ca_RPMI_DOWNREG<-results[which((results$logFC<(-FC_thresh)&results$adj.P.Val<adj_p_value_thresh)),]
CaKp_vs_Ca_RPMI_DEGS=rownames(results_CaKp_vs_Ca_RPMI_DEGS)
CaKp_vs_Ca_RPMI_DEGS_UPREG=rownames(results_CaKp_vs_Ca_RPMI_UPREG)
CaKp_vs_Ca_RPMI_DEGS_DOWNREG=rownames(results_CaKp_vs_Ca_RPMI_DOWNREG)
add_attributes<-c("symbol","name",colnames(dataframe_expression_table)[grep("C",colnames(dataframe_expression_table))])
results_CaKp_vs_Ca_RPMI_DEGS[,add_attributes]<-dataframe_expression_table[match(CaKp_vs_Ca_RPMI_DEGS,dataframe_expression_table[,"locus_tag"]),add_attributes]
create_dt(results_CaKp_vs_Ca_RPMI_DEGS)
save_xlsx(results_CaKp_vs_Ca_RPMI_DEGS)
```

__Summary table of Up/Down/NotSig__
```{r ca summary degs cakp vs ca rpmi,echo=FALSE}
summary_table<-summary(decideTests(fit_2,lfc=FC_thresh,p.value=adj_p_value_thresh))
colnames(summary_table)<-gsub("=.*","",colnames(summary_table))
summary_table
```
__Volcano plots__
```{r ca volcanos cakp vs ca rpmi,fig.height=7,fig.width=12, out.width = "\\textwidth*10", out.height = "\\textheight*800"}
top_n=40
res<-results
res$symbol<-dataframe_expression_table[rownames(results),"symbol"]
res_sig<-res[res$adj.P.Val<adj_p_value_thresh,]
sig_down<-res_sig[which(res_sig$logFC<(-FC_thresh)),]
top_sig_down<-head(sig_down$symbol[sig_down$symbol!=""],top_n)
fc_down<-sig_down[order(sig_down$logFC),]
top_fc_down<-head(fc_down$symbol[fc_down$symbol!=""],top_n)
sig_up<-res_sig[which(res$logFC>(FC_thresh)),]
top_sig_up<-head(sig_up$symbol[sig_up$symbol!=""],top_n)
fc_up<-sig_up[order(sig_up$logFC,decreasing=TRUE),]
top_fc_up<-head(fc_up$symbol[fc_up$symbol!=""],top_n)
genes_of_interest<-c(top_sig_down,top_sig_up,top_fc_down,top_fc_up)
p<-plotVolcano(results_obj = res,genes_of_interest = genes_of_interest,plot_title=colnames(contrast.matrix),FC_thresh=log2(FC),adj_p_value_thresh=adj_p_value_thresh)
vplots<-list()
vplots[[length(vplots)+1]]<-p
pdf(file.path(OUTPUT_DIR,"CaKp_vs_Ca_RPMI_volcanoplot.pdf"),height=8,width=8)
p
dev.off()
p
```

__Volcano plots: Gene list labels__
```{r ca volcanos cakp vs ca rpmi label,fig.height=7,fig.width=12, out.width = "\\textwidth*10", out.height = "\\textheight*800"}
label_up<-c("YWP1", "DAG7", "ALS1", "ALS2", "ALS4", "PCK1", "FBP1", "ICL1", "MDH1", "CYP1", "FOX3", "CAT2", "CEK2", "SIM1", "HSP21", "HSP104")
label_down<-c("HYR1", "PRA1", "ZRT1", "CRZ2", "CHS1", "CHT4", "PGA53", "PGA50", "ALS3", "SAP5")
genes_of_interest<-c(label_up,label_down)
p<-plotVolcano2(results_obj = res,genes_of_interest = genes_of_interest,plot_title=colnames(contrast.matrix),FC_thresh=log2(FC),adj_p_value_thresh=adj_p_value_thresh)
vplots<-list()
vplots[[length(vplots)+1]]<-p
pdf(file.path(OUTPUT_DIR,"CaKp_vs_Ca_RPMI_volcanoplot_genes_of_interest.pdf"),height=6,width=8)
p
dev.off()
p
```

##### __Heatmap of gene expression for RPMI samples (Top Upregulated and Downregulated DEGs)__

```{r ca top deg heatmap RPMI}
top_n=50
top_fc_down<-head(fc_down,top_n)
top_fc_up<-head(fc_up,top_n)
top_fc_genes<-c(rownames(top_fc_down),rownames(top_fc_up))
samples_order<-c(metadata[which(metadata[,"group_name_1"]=="RPMI" & metadata[,"group_name_2"]=="Ca mix/Kp mix"),]$sample_name_in_report,metadata[which(metadata[,"group_name_1"]=="RPMI" & metadata[,"group_name_2"]=="Ca single"),]$sample_name_in_report)
lcpm_heatmap<-lcpm[top_fc_genes,samples_order]
rownames(lcpm_heatmap)<-c(paste(rownames(top_fc_down),top_fc_down$symbol,sep="-"),paste(rownames(top_fc_up),top_fc_up$symbol,sep="-"))
p<-pheatmap(lcpm_heatmap, scale = "row",show_rownames = TRUE,heatmap_legend_param = list(title = "Row Z-score"),name="RPMI",col=rev(brewer.pal(n = 8, name = "RdBu")),cellwidth=8,cellheight=8, fontsize=8, fontsize_row=6, cluster_cols = FALSE)
pdf(file.path(OUTPUT_DIR,"Ca_top_DEGs_expression_RPMI_heatmap.pdf"),height=13,width=5)
p
dev.off()
p

```

### __Comparison: C. Cakp vs Ca (from M9 group)__

#### __Contrast matrix__
```{r ca Contrast cakp vs ca m9}
contrast.matrix <- makeContrasts(CaKp_vs_Ca_M9 = CaKp_mix__M9 - Ca_single__M9, levels=colnames(design))
```

#### __Significant DEGs__
```{r ca DEG cakp vs ca m9}
fit_2 <- contrasts.fit(fit,contrast.matrix)
fit_2 <- eBayes(fit_2)

##Benjamini-Hochberg method used to adjust the p-values for multiple testing
results<-topTable(fit_2, adjust="BH",number=nrow(filtered_y))
results_CaKp_vs_Ca_M9<-results
#all significant DEGs based on p-value and FC thresholds
results_CaKp_vs_Ca_M9_DEGS<-results[which((abs(results$logFC)>FC_thresh&results$adj.P.Val<adj_p_value_thresh)),]
results_CaKp_vs_Ca_M9_UPREG<-results[which((results$logFC>FC_thresh&results$adj.P.Val<adj_p_value_thresh)),]
results_CaKp_vs_Ca_M9_DOWNREG<-results[which((results$logFC<(-FC_thresh)&results$adj.P.Val<adj_p_value_thresh)),]
CaKp_vs_Ca_M9_DEGS=rownames(results_CaKp_vs_Ca_M9_DEGS)
CaKp_vs_Ca_M9_DEGS_UPREG=rownames(results_CaKp_vs_Ca_M9_UPREG)
CaKp_vs_Ca_M9_DEGS_DOWNREG=rownames(results_CaKp_vs_Ca_M9_DOWNREG)
add_attributes<-c("symbol","name",colnames(dataframe_expression_table)[grep("C",colnames(dataframe_expression_table))])
results_CaKp_vs_Ca_M9_DEGS[,add_attributes]<-dataframe_expression_table[match(CaKp_vs_Ca_M9_DEGS,dataframe_expression_table[,"locus_tag"]),add_attributes]
create_dt(results_CaKp_vs_Ca_M9_DEGS)
save_xlsx(results_CaKp_vs_Ca_M9_DEGS)
```

__Summary table of Up/Down/NotSig__
```{r ca summary degs tables cakp vs ca m9,echo=FALSE}
summary_table<-summary(decideTests(fit_2,lfc=FC_thresh,p.value=adj_p_value_thresh))
colnames(summary_table)<-gsub("=.*","",colnames(summary_table))
summary_table
```

__Volcano plots__
```{r ca volcanos cakp vs ca m9,fig.height=7,fig.width=12, out.width = "\\textwidth*10", out.height = "\\textheight*800"}
res<-results
res$symbol<-dataframe_expression_table[rownames(results),"symbol"]
res_sig<-res[res$adj.P.Val<adj_p_value_thresh,]
sig_down<-res_sig[which(res_sig$logFC<(-FC_thresh)),]
top_sig_down<-head(sig_down$symbol[sig_down$symbol!=""],top_n)
fc_down<-sig_down[order(sig_down$logFC),]
top_fc_down<-head(fc_down$symbol[fc_down$symbol!=""],top_n)
sig_up<-res_sig[which(res$logFC>(FC_thresh)),]
top_sig_up<-head(sig_up$symbol[sig_up$symbol!=""],top_n)
fc_up<-sig_up[order(sig_up$logFC,decreasing=TRUE),]
top_fc_up<-head(fc_up$symbol[fc_up$symbol!=""],top_n)
genes_of_interest<-c(top_sig_down,top_sig_up,top_fc_down,top_fc_up)
p<-plotVolcano(results_obj = res,genes_of_interest = genes_of_interest,plot_title=colnames(contrast.matrix),FC_thresh=log2(FC),adj_p_value_thresh=adj_p_value_thresh)
vplots<-list()
vplots[[length(vplots)+1]]<-p
pdf(file.path(OUTPUT_DIR,"CaKp_vs_Ca_M9_volcanoplot.pdf"),height=8,width=8)
p
dev.off()
p
```

__Volcano plots: Gene list labels__
```{r ca volcanos cakp vs ca m9 label,fig.height=7,fig.width=12, out.width = "\\textwidth*10", out.height = "\\textheight*800"}
label_up<-c("PCK1", "FBP1", "CIT1", "ACS1", "TRY4", "TRY6", "PRA1", "ZRT1", "ZRT2", "PHO84", "PHO86", "HGT16", "PRY1", "CTN1", "SAP6", "MAL31", "LAP3", "CSA2")
label_down<-c("PGA11", "PGA23", "PGA30", "PGA58", "PGA62", "PGA63", "MNN1", "XOG1", "CHS1", "CHS2", "SAP7")
genes_of_interest<-c(label_up,label_down)
p<-plotVolcano2(results_obj = res,genes_of_interest = genes_of_interest,plot_title=colnames(contrast.matrix),FC_thresh=log2(FC),adj_p_value_thresh=adj_p_value_thresh)
vplots<-list()
vplots[[length(vplots)+1]]<-p
pdf(file.path(OUTPUT_DIR,"CaKp_vs_Ca_M9_volcanoplot_genes_of_interest.pdf"),height=6,width=8)
p
dev.off()
p
```

#### __Heatmap of gene expression for M9 samples (Top Upregulated and Downregulated DEGs)__

```{r ca top deg heatmap M9}
top_n=50
top_fc_down<-head(fc_down,top_n)
top_fc_up<-head(fc_up,top_n)
top_fc_genes<-c(rownames(top_fc_down),rownames(top_fc_up))
samples_order<-c(gsub(" ", "",metadata[which(metadata[,"group_name_1"]=="M9" & metadata[,"group_name_2"]=="Ca mix/Kp mix"),]$sample_name_in_report),gsub(" ", "",metadata[which(metadata[,"group_name_1"]=="M9" & metadata[,"group_name_2"]=="Ca single"),]$sample_name_in_report))
lcpm_heatmap<-lcpm[top_fc_genes,samples_order]
rownames(lcpm_heatmap)<-c(paste(rownames(top_fc_down),top_fc_down$symbol,sep="-"),paste(rownames(top_fc_up),top_fc_up$symbol,sep="-"))
p<-pheatmap(lcpm_heatmap, scale = "row",show_rownames = TRUE,heatmap_legend_param = list(title = "Row Z-score"),name="M9",col=rev(brewer.pal(n = 8, name = "RdBu")),cellwidth=8,cellheight=8, fontsize=8, fontsize_row=6, cluster_cols = FALSE)
pdf(file.path(OUTPUT_DIR,"Ca_top_DEGs_expression_M9_heatmap.pdf"),height=13,width=5)
p
dev.off()
p
```

#### __CaKp vs Ca DEGs venn__

```{r overlap comp cakp vs ca ,echo=FALSE,fig.height=7,fig.width=12,out.width='85%'}
signatures<-list(CaKp_vs_Ca_RPMI_DEGS_UPREG=CaKp_vs_Ca_RPMI_DEGS_UPREG,CaKp_vs_Ca_RPMI_DEGS_DOWNREG=CaKp_vs_Ca_RPMI_DEGS_DOWNREG,CaKp_vs_Ca_M9_DEGS_UPREG=CaKp_vs_Ca_M9_DEGS_UPREG,CaKp_vs_Ca_M9_DEGS_DOWNREG=CaKp_vs_Ca_M9_DEGS_DOWNREG)

library(eulerr)
vennfit <- euler(signatures)
plot(vennfit,
     fills = c(brewer.pal(n = 8, name = "RdBu")[2],brewer.pal(n = 8, name = "RdBu")[7],brewer.pal(n = 8, name = "RdBu")[3], brewer.pal(n = 8, name = "RdBu")[6]),
     edges = TRUE,
     fontsize = 8,
     quantities = list(fontsize = 8))

```

```{r ca gene lists }
write_venn_genelists(signatures = signatures,sig_names_interest = names(signatures), outbasedir=file.path(OUTPUT_DIR,"CaKp_vs_Ca__venndiagram_genes"))
```

### __GO Enrichment analysis with TopGO__ 

In this section the R package   [TopGO](https://bioconductor.org/packages/release/bioc/html/topGO.html "Alexa A, Rahnenfuhrer J (2022). topGO: Enrichment Analysis for Gene Ontology. R package version 2.49.0.") was used to reanalyse GO Biological Process terms, accounting for the relationship between parent and child GO terms. The algorithm used was `weight01`. Statistical significance was tested with the Fisher's exact test and multiple testing correction was performed using the Benjamini-Hochberg method.


```{r ca topGO setup}
#GO term look up
gomappingsfile="GCF_000182965.3_ASM18296v3_gene_ontology_geneid2go_mappings.tsv"
geneID2GO<-readMappings(file=file.path(ANNOTATION_DIR, gomappingsfile))
gotable<-read.table(file.path(ANNOTATION_DIR, gomappingsfile))
colnames(gotable)<-c("locus_tag","go_terms")
create_dt(gotable)
```


#### __RPMI__

```{r ca topGO degs RPMI}
media="RPMI"
#1. Run TopGO for each signature
geneNames<-names(geneID2GO)
GOdata_list<-lapply(grep(media,names(signatures),value = T),function(comp){
  interestingGenes<-signatures[[comp]]
  geneList<-factor(as.integer(geneNames %in% interestingGenes))
  names(geneList)<-geneNames
  GOdata <- new('topGOdata',
                ontology = 'BP',
                allGenes = geneList,
                annot = annFUN.gene2GO,
                gene2GO = geneID2GO,
                nodeSize = 10)
  return(GOdata)  
})
names(GOdata_list)<-grep(media,names(signatures),value = T)

```

```{r ca significant terms rpmi}

#2. Apply Fisher exact test with weight01 algorithm from TopGO
test.stat <- new("weight01Count", testStatistic = GOFisherTest, name = "Fisher test")

#3.-7. Run topGO_wrapper that produces a list of tables for each comparison for all significant terms and a comparison plot 
topGO_results<-topGO_wrapper(GOdata_list,test.stat,FDR_thres=0.05)

# Table of genes in each GO term
topGO_results<-genes_per_pathway(GOdata_list,signatures,topGO_results=topGO_results)

##Write file as excel file with multiple worksheets
openxlsx::write.xlsx(topGO_results$result_list,file.path(OUTPUT_DIR,paste(media,"ca_degs_topGO_results.xlsx",sep="_")),colNames=TRUE,rowNames=FALSE)
```

```{r ca topgo plot rpmi,echo=FALSE,fig.width=8,fig.height=6+length(unique(topGO_results$sigplot$data$Term))/4}
topGO_results$sigplot
```

#### __M9__

```{r ca topGO degs M9}
media="M9"
#1. Run TopGO for each signature
geneNames<-names(geneID2GO)
GOdata_list<-lapply(grep(media,names(signatures),value = T),function(comp){
  interestingGenes<-signatures[[comp]]
  geneList<-factor(as.integer(geneNames %in% interestingGenes))
  names(geneList)<-geneNames
  GOdata <- new('topGOdata',
                ontology = 'BP',
                allGenes = geneList,
                annot = annFUN.gene2GO,
                gene2GO = geneID2GO,
                nodeSize = 10)
  return(GOdata)  
})
names(GOdata_list)<-grep(media,names(signatures),value = T)

```

```{r ca significant terms m9}
#2. Apply Fisher exact test with weight01 algorithm from TopGO
test.stat <- new("weight01Count", testStatistic = GOFisherTest, name = "Fisher test")

#3.-7. Run topGO_wrapper that produces a list of tables for each comparison for all significant terms and a comparison plot 
topGO_results<-topGO_wrapper(GOdata_list[grep(media,names(GOdata_list))],test.stat,FDR_thres=0.05)

# Table of genes in each GO term
topGO_results<-genes_per_pathway(GOdata_list[grep(media,names(GOdata_list))],signatures[grep(media,names(signatures))],topGO_results=topGO_results)

##Write file as excel file with multiple worksheets
openxlsx::write.xlsx(topGO_results$result_list,file.path(OUTPUT_DIR,paste(media,"ca_degs_topGO_results.xlsx",sep="_")),colNames=TRUE,rowNames=FALSE)
```

```{r ca topgo plot media,echo=FALSE,fig.width=8,fig.height=6+length(unique(topGO_results$sigplot$data$Term))/4}
topGO_results$sigplot
```

### __KEGG Pathways analysis: Over Representation/Enrichment analysis__

The KEGG over-representation, GSEA and `pathview` analyses in this report, for both *C. albicans* (`organism = "cal"`) and *K. pneumoniae* (`organism = "kpn"`), were produced with [clusterProfiler](https://bioconductor.org/packages/release/bioc/html/clusterProfiler.html) and [pathview](https://bioconductor.org/packages/release/bioc/html/pathview.html), using the Benjamini-Hochberg method for multiple testing correction.

**Reproducibility note on KEGG:** KEGG pathway and gene mappings are retrieved using the KEGG REST API when the report is generated. The outputs from _enrichKEGG_, _gseKEGG_, _pathview_ and _keggGet_ reflect the state of the cal (_C. albicans_) and kpn (_K. pneumoniae_) databases on the date this report was rendered. This may result in slight differences in pathway memberships, gene counts, and p-values from the published version.

#### __All DEGs RPMI__
```{r ora ca rpmi}
keggEnr<-enrichKEGG(rownames(results_CaKp_vs_Ca_RPMI_DEGS),organism='cal',pvalueCutoff = 0.05)
keggEnr_RPMI<-keggEnr
create_dt(keggEnr@result)
```
```{r ora ca rpmi plot,fig.width=10,fig.height=(4+dim(keggEnr)[1]/2)}
dotplot_scale(keggEnr, title="All DEGs RPMI",font.size=10)
```

#### __Upregulated DEGs RPMI__
```{r ora ca rpmi up}
keggEnr<-enrichKEGG(rownames(results_CaKp_vs_Ca_RPMI_UPREG),organism='cal',pvalueCutoff = 0.05)
create_dt(keggEnr@result)
```

```{r ora ca rpmi up plot,fig.width=10,fig.height=(4+dim(keggEnr)[1]/2)}
dotplot_scale(keggEnr, title="Upreg DEGs RPMI",font.size=10)
```

#### __Downregulated DEGs RPMI__ 
```{r ora ca rpmi down}
keggEnr<-enrichKEGG(rownames(results_CaKp_vs_Ca_RPMI_DOWNREG),organism='cal',pvalueCutoff = 0.05)
create_dt(keggEnr@result)
```

```{r ora ca rpmi down plot,fig.width=10,fig.height=(4+dim(keggEnr)[1]/2)}
dotplot_scale(keggEnr, title="Downreg DEGs RPMI",font.size=10)
```

#### __All DEGs M9__ 
```{r ora ca m9}
keggEnr<-enrichKEGG(rownames(results_CaKp_vs_Ca_M9_DEGS),organism='cal',pvalueCutoff = 0.05)
keggEnr_M9<-keggEnr
create_dt(keggEnr@result)
```

```{r ora ca m9 plot,fig.width=10,fig.height=(4+dim(keggEnr)[1]/2)}
dotplot_scale(keggEnr, title="All DEGs M9",font.size=10)
```

#### __Upregulated DEGs RPMI__
```{r ora ca m9 up}
keggEnr<-enrichKEGG(rownames(results_CaKp_vs_Ca_M9_UPREG),organism='cal',pvalueCutoff = 0.05)
create_dt(keggEnr@result)
```

```{r ora ca m9 up plot,fig.width=10,fig.height=(4+dim(keggEnr)[1]/2)}
dotplot_scale(keggEnr, title="Upreg DEGs M9",font.size=10)
```

#### __Downregulated DEGs RPMI__ 
```{r ora ca m9 down}
keggEnr<-enrichKEGG(rownames(results_CaKp_vs_Ca_M9_DOWNREG),organism='cal',pvalueCutoff = 0.05)
create_dt(keggEnr@result)
```

```{r ora ca m9 down plot,fig.width=10,fig.height=(4+dim(keggEnr)[1]/2)}
dotplot_scale(keggEnr, title="Downreg DEGs M9",font.size=10)
```

### __Heatmaps for KEGG Over Representation/Enrichment analysis__

#### __Combined significant KEGG pathways in both media__

>IMPT only significantly differentially expressed genes are included in these heatmaps as some KEGG pathways have over a thousand genes

```{r signif kegg}
keggEnr_RPMI_signif<-keggEnr_RPMI@result[which(keggEnr_RPMI@result$p.adjust<0.05),'ID']
keggEnr_M9_signif<-keggEnr_M9@result[which(keggEnr_M9@result$p.adjust<0.05),'ID']
signif_kegg_pathways<-unique(c(keggEnr_RPMI_signif,keggEnr_M9_signif))
```

```{r heatmap for each pathway}
for(keggpathway in signif_kegg_pathways){
  all_signif_genes_in_keggpathway<-c()
  if(keggpathway %in% keggEnr_RPMI[,'ID']){
    title=keggEnr_RPMI[keggpathway,'Description']
    all_signif_genes_in_keggpathway<-c(all_signif_genes_in_keggpathway,unlist(strsplit(keggEnr_RPMI[keggpathway,'geneID'],"/")))
  }
  if(keggpathway %in% keggEnr_M9[,'ID']){
    title=keggEnr_M9[keggpathway,'Description']
    all_signif_genes_in_keggpathway<-c(all_signif_genes_in_keggpathway,unlist(strsplit(keggEnr_M9[keggpathway,'geneID'],"/")))
  }
  
  #identify genes' logFC and p-value for each medium
  CaKp_vs_Ca_RPMI_log2FC<-vector("numeric",length=length(all_signif_genes_in_keggpathway))
  CaKp_vs_Ca_M9_log2FC<-vector("numeric",length=length(all_signif_genes_in_keggpathway))
  CaKp_vs_Ca_RPMI_adjpval<-vector("numeric",length=length(all_signif_genes_in_keggpathway))
  CaKp_vs_Ca_M9_adjpval<-vector("numeric",length=length(all_signif_genes_in_keggpathway))  
  for (i in c(1:length(all_signif_genes_in_keggpathway))){
    CaKp_vs_Ca_RPMI_log2FC[i]<-results_CaKp_vs_Ca_RPMI[all_signif_genes_in_keggpathway[i],'logFC']
    CaKp_vs_Ca_M9_log2FC[i]<-results_CaKp_vs_Ca_M9[all_signif_genes_in_keggpathway[i],'logFC']
    CaKp_vs_Ca_RPMI_adjpval[i]<-results_CaKp_vs_Ca_RPMI[all_signif_genes_in_keggpathway[i],'adj.P.Val']
    CaKp_vs_Ca_M9_adjpval[i]<-results_CaKp_vs_Ca_M9[all_signif_genes_in_keggpathway[i],'adj.P.Val']
  }
  matlog2fc<-cbind(CaKp_vs_Ca_RPMI_log2FC,CaKp_vs_Ca_M9_log2FC)
  rownames(matlog2fc)<-paste(all_signif_genes_in_keggpathway,dataframe_expression_table[all_signif_genes_in_keggpathway,"symbol"],sep="-")
  colnames(matlog2fc)<-gsub("_log2FC","",colnames(matlog2fc))
  matadjpval<-cbind(CaKp_vs_Ca_RPMI_adjpval,CaKp_vs_Ca_M9_adjpval)

  for (j in seq_len(ncol(matlog2fc))) {
  # Extract the j-th column from both matrices
  mat_col <- matrix(matlog2fc[, j], ncol = 1)
  rownames(mat_col) <- rownames(matlog2fc)
  colnames(mat_col) <- colnames(matlog2fc)[j] 

  mat_pval_col <- matrix(matadjpval[, j], ncol = 1)
  rownames(mat_pval_col) <- rownames(matadjpval)
  colnames(mat_pval_col) <- colnames(matlog2fc)[j]  # align for cell_fun logic

  # Create the heatmap
  H <- Gene_Heatmap(mat_col,mat_pval_col,title)

  # Set height dynamically
  nr <- nrow(matlog2fc)
  pdf(
    file.path(OUTPUT_DIR, paste("KEGG_ORA", keggpathway, colnames(matlog2fc)[j], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
  )
  draw(H)
  dev.off()
}


}

```


Note that plots have been saved as PDF files and are not included in the report

### __KEGG Pathways analysis: Gene Set Enrichment Analysis (GSEA)__

See the KEGG reproducibility note in the over-representation section above.

Note only the top 10 pathways are shown in the figures.

#### __RPMI__ 

```{r gsea ca rpmi}
geneList<-as.matrix(results_CaKp_vs_Ca_RPMI$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_RPMI)
keggGs<-gseKEGG(geneList = sort(geneList[,1],decreasing = T), organism = 'cal', minGSSize = 10, pvalueCutoff = 0.05, verbose = F,eps =0)
create_dt(keggGs@result)

```

```{r gsea ca rpmi plot hide, fig.height=10,fig.width=10,results='hide',echo=FALSE,fig.show='hide'}
final_plot<-modgseaplot(keggGs,geneSetID,colourlist)
```

```{r gsea ca rpmi plot, fig.height=4,fig.width=13,echo=FALSE,out.width = "200%"}
print(final_plot)
```


#### __M9__ 

```{r gsea ca m9}
geneList<-as.matrix(results_CaKp_vs_Ca_M9$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_M9)
keggGs<-gseKEGG(geneList = sort(geneList[,1],decreasing = T), organism = 'cal', minGSSize = 10, pvalueCutoff = 0.05, verbose = F,eps =0)
create_dt(keggGs@result)
```


```{r gsea ca m9 plot hide, fig.height=10,fig.width=10,results='hide',echo=FALSE,fig.show='hide'}
final_plot<-modgseaplot(keggGs,geneSetID,colourlist)
```

```{r gsea ca m9 plot, fig.height=4,fig.width=14,echo=FALSE,out.width = "200%"}
print(final_plot)
```



## __Candida Biofilm Transcription Factor expression__

__Expression of list of TFs from Tamires:__
```{r tf genes from Tamires}
tf_file<-"TF_biofilm_Ca.txt"
tfs <- read.table(tf_file,header=F)
tf_gene_info<-gene_info[which(gene_info$symbol %in% toupper(tfs[,1])),]
tf_exp<-cbind(tf_gene_info,dataframe_expression_table[tf_gene_info$locus_tag,])
create_dt(tf_exp)
```

__Log counts per million transform for visualization:__
```{r lcpm}
tf_lcpm<-cbind(tf_gene_info,lcpm[tf_gene_info$locus_tag,])
```
__Z scaling:__
```{r tf expression data}
tf_RPMI_data<-lcpm[tf_gene_info$locus_tag,grep("n",colnames(lcpm),value=T,invert = T)]
tf_M9_data<-lcpm[tf_gene_info$locus_tag,grep("n",colnames(lcpm),value=T)]
Z.tf_RPMI_data<-t(scale(t(tf_RPMI_data)))
Z.tf_M9_data<-t(scale(t(tf_M9_data)))
```

__Combined heatmaps__
```{r heatmap target genes,fig.height=6,fig.width=8}
H1<-Heatmap(Z.tf_RPMI_data,name='Row Z-score',col = rev(brewer.pal(n = 7, name = "RdBu")), row_labels=tf_gene_info$symbol, cluster_rows = F,column_title = 'RPMI',column_names_side = 'top')
H2<-Heatmap(Z.tf_M9_data,name='Row Z-score',col = rev(brewer.pal(n = 7, name = "RdBu")), row_labels=tf_gene_info$symbol, cluster_rows = F,column_title = 'M9',column_names_side = 'top')
Hlist<-H1+H2
draw(Hlist, row_title = "Ca Biofilm TFs")
```
__RPMI__
```{r heatmap target genes RPMI,fig.height=6,fig.width=5,out.width="50%"}
draw(H1, row_title = "Ca Biofilm TFs")
```

__M9__
```{r heatmap target genes M9,fig.height=6,fig.width=5,out.width="50%"}
draw(H2, row_title = "Ca Biofilm TFs")
```


## __Pathways of interest: RPMI__

The `pathview` maps and `keggGet` pathway descriptions below are retrieved from KEGG; see the KEGG reproducibility note in the over-representation section above.

```{r setup pathways rpmi, echo=FALSE}
#variables for filename
spec="ca"
media="rpmi"

#Prepare table with gene name, fc and adj pvalue + add expression of samples
subset_results<-cbind(results_CaKp_vs_Ca_RPMI$logFC,results_CaKp_vs_Ca_RPMI$adj.P.Val)
rownames(subset_results)<-rownames(results_CaKp_vs_Ca_RPMI)
colnames(subset_results)<-c("logFC","adj.P.Val")

info_cols<-ca_counts_per_million_expression_table[rownames(subset_results),c("locus_tag","symbol","name")]
rowsorted_expression_data<-ca_counts_per_million_expression_table[rownames(subset_results),c(4:dim(ca_counts_per_million_expression_table)[2])]
full_data<-cbind(info_cols,subset_results,rowsorted_expression_data)

pathview_low<-list(gene=brewer.pal(n = 8, name = "RdBu")[8],cpd=brewer.pal(n = 8, name = "RdBu")[8])
pathview_mid<-list(gene="white",cpd="white")
pathview_high<-list(gene=brewer.pal(n = 8, name = "RdBu")[1],cpd=brewer.pal(n = 8, name = "RdBu")[1])
```

### __1. UPR__

```{r, echo=FALSE}
pathid="cal04141"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_RPMI$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_RPMI)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))),cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_RPMI"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_RPMI"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __2. ROS__

cal04213 is no longer in KEGG

  
### __3. MAPK__
  
```{r, echo=FALSE}
pathid="cal04011"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_RPMI$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_RPMI)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_RPMI"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_RPMI"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __4. Ergosterol__

```{r, echo=FALSE}
pathid="cal00100"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_RPMI$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_RPMI)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_RPMI"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_RPMI"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __5. Fatty acid degradation__

```{r, echo=FALSE}
pathid="cal00071"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_RPMI$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_RPMI)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_RPMI"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_RPMI"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __6. Glyoxylate__

```{r, echo=FALSE}
pathid="cal00630"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_RPMI$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_RPMI)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_RPMI"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_RPMI"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __7. Cell cycle__

```{r, echo=FALSE}
pathid="cal04111"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_RPMI$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_RPMI)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```


__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_RPMI"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_RPMI"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __8. TCA__

```{r, echo=FALSE}
pathid="cal00020"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_RPMI$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_RPMI)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_RPMI"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_RPMI"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __9. Cofactors__

__No genes could be retrieved from KEGG for this pathway (cal01240)!__

  
### __10. Porhyrins__
  
```{r, echo=FALSE}
pathid="cal00860"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_RPMI$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_RPMI)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_RPMI"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_RPMI"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __11. NON KEGG PATHWAY:  Hyphal Growth Genes__

__Converted list of genes from Tamires:__

```{r hyphal genes from Tamires}
custom_genes <- read.table("hyphal_genes.txt",header=F)
custom_genes_info<-gene_info[which(gene_info$locus_tag %in% custom_genes[,1]),]
custom_genes_exp<-cbind(custom_genes_info,dataframe_expression_table[custom_genes_info$locus_tag,])
create_dt(custom_genes_exp)
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE,message=FALSE}
all_genes_in_pathway<-custom_genes[,1]
matlog2fc <- matrix(nrow=length(all_genes_in_pathway),ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_pathway),ncol=1)
for (i in c(1:length(all_genes_in_pathway))){
  matlog2fc[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_pathway[i],'logFC'])
  matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_pathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_pathway,dataframe_expression_table[all_genes_in_pathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_RPMI"
rownames(matadjpval)<-paste(all_genes_in_pathway,dataframe_expression_table[all_genes_in_pathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_RPMI"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title="Hyphal Growth Genes")
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
pathway="hyphal_growth_genes"
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("NON_KEGG_Requested", pathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __12. NON KEGG PATHWAY:  CWI Pathway Genes__

__Converted list of genes from Tamires:__

```{r cwi genes from Tamires RPMI}
custom_genes <- read.table("cwi_genes.txt",header=F)
custom_genes_info<-gene_info[which(gene_info$locus_tag %in% custom_genes[,1]),]
custom_genes_exp<-cbind(custom_genes_info,dataframe_expression_table[custom_genes_info$locus_tag,])
create_dt(custom_genes_exp)
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_pathway<-custom_genes[,1]
matlog2fc <- matrix(nrow=length(all_genes_in_pathway),ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_pathway),ncol=1)
for (i in c(1:length(all_genes_in_pathway))){
  matlog2fc[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_pathway[i],'logFC'])
  matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_pathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_pathway,dataframe_expression_table[all_genes_in_pathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_RPMI"
rownames(matadjpval)<-paste(all_genes_in_pathway,dataframe_expression_table[all_genes_in_pathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_RPMI"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title="CWI Pathway Genes")
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
pathway="cwi_pathway_genes"
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("NON_KEGG_Requested", pathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```


### __13. NON KEGG PATHWAY:  RAM Pathway Genes__

__Converted list of genes from Tamires:__

```{r ram genes from Tamires RPMI}
custom_genes <- read.table("ram_genes.txt",header=F)
custom_genes_info<-gene_info[which(gene_info$locus_tag %in% custom_genes[,1]),]
custom_genes_exp<-cbind(custom_genes_info,dataframe_expression_table[custom_genes_info$locus_tag,])
create_dt(custom_genes_exp)
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_pathway<-custom_genes[,1]
matlog2fc <- matrix(nrow=length(all_genes_in_pathway),ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_pathway),ncol=1)
for (i in c(1:length(all_genes_in_pathway))){
  matlog2fc[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_pathway[i],'logFC'])
  matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_RPMI[all_genes_in_pathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_pathway,dataframe_expression_table[all_genes_in_pathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_RPMI"
rownames(matadjpval)<-paste(all_genes_in_pathway,dataframe_expression_table[all_genes_in_pathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_RPMI"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title="RAM Pathway Genes")
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
pathway="ram_pathway_genes"
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("NON_KEGG_Requested", pathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```


## __Pathways of interest: M9__

The `pathview` maps and `keggGet` pathway descriptions below are retrieved from KEGG; see the KEGG reproducibility note in the over-representation section above.

```{r setup pathways M9, echo=FALSE}
#variables for filename
spec="ca"
media="m9"

#Prepare table with gene name, fc and adj pvalue + add expression of samples
subset_results<-cbind(results_CaKp_vs_Ca_M9$logFC,results_CaKp_vs_Ca_M9$adj.P.Val)
rownames(subset_results)<-rownames(results_CaKp_vs_Ca_M9)
colnames(subset_results)<-c("logFC","adj.P.Val")

info_cols<-ca_counts_per_million_expression_table[rownames(subset_results),c("locus_tag","symbol","name")]
rowsorted_expression_data<-ca_counts_per_million_expression_table[rownames(subset_results),c(4:dim(ca_counts_per_million_expression_table)[2])]
full_data<-cbind(info_cols,subset_results,rowsorted_expression_data)
```

### __1. UPR__

```{r, echo=FALSE}
pathid="cal04141"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_M9$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_M9)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_M9"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_M9"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __2. ROS__

cal04213 is no longer in KEGG


### __3. MAPK__
  
```{r, echo=FALSE}
pathid="cal04011"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_M9$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_M9)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_M9"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_M9"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __4. Ergosterol__

```{r, echo=FALSE}
pathid="cal00100"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_M9$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_M9)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_M9"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_M9"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __5. Fatty acid degradation__

```{r, echo=FALSE}
pathid="cal00071"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_M9$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_M9)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1))
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_M9"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_M9"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __6. Glyoxylate__

```{r, echo=FALSE}
pathid="cal00630"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_M9$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_M9)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_M9"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_M9"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __7. Cell cycle__

```{r, echo=FALSE}
pathid="cal04111"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_M9$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_M9)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_M9"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_M9"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __8. TCA__

```{r, echo=FALSE}
pathid="cal00020"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_M9$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_M9)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_M9"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_M9"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __9. Cofactors__

__No genes could be retrieved from KEGG for this pathway (cal01240)!__
  
### __10. Porhyrins__
  
```{r, echo=FALSE}
pathid="cal00860"
res<-keggGet(pathid)

geneList<-as.matrix(results_CaKp_vs_Ca_M9$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Ca_M9)
genelogFC<-geneList[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]
names(genelogFC)<-rownames(geneList)[which(rownames(geneList) %in% res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])]

genelogFC<-genelogFC[!is.na(genelogFC)]

if (length(genelogFC)>0){
  pathviewres<-pathview(gene.data = genelogFC, 
                        gene.idtype = "kegg",
                        pathway.id = pathid,
                        kegg.dir = OUTPUT_DIR,
                        species='cal',
                        limit  = list(gene=round(max(abs(genelogFC))), cpd=1),
                        low = pathview_low,
                        mid = pathview_mid,
                        high = pathview_high)
}
```


Of `r length(res[[1]]$GENE[seq(1,length(res[[1]]$GENE),2)])` genes in the KEGG pathway (id `r pathid`), `r length(genelogFC)` gene(s) were found in the gene expression data.

```{r, echo =FALSE}
if (length(genelogFC)>0){
  nf=file.path(OUTPUT_DIR,paste0(pathid,".pathview.",spec,".",media,".png"))
  file.rename(paste0("./",pathid,".pathview.png"),nf)
}
```

`r sprintf("![Fold change of all genes associated to the KEGG pathway](%s)", nf)`

__List of matched genes in pathway__
```{r, echo=FALSE}
create_dt(full_data[rownames(full_data) %in% names(genelogFC),])
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_keggpathway<-names(genelogFC)
matlog2fc <- matrix(genelogFC, ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_keggpathway),ncol=1)
for (i in c(1:length(all_genes_in_keggpathway))){
    matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_keggpathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_M9"
rownames(matadjpval)<-paste(all_genes_in_keggpathway,dataframe_expression_table[all_genes_in_keggpathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_M9"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title=res[[1]]$NAME)
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("KEGG_Requested", keggpathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```



### __11. NON KEGG PATHWAY:  Hyphal Growth Genes__

__Converted list of genes from Tamires:__

```{r hyphal genes from Tamires m9}
custom_genes <- read.table("hyphal_genes.txt",header=F)
custom_genes_info<-gene_info[which(gene_info$locus_tag %in% custom_genes[,1]),]
custom_genes_exp<-cbind(custom_genes_info,dataframe_expression_table[custom_genes_info$locus_tag,])
create_dt(custom_genes_exp)
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE,message=FALSE}
all_genes_in_pathway<-custom_genes[,1]
matlog2fc <- matrix(nrow=length(all_genes_in_pathway),ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_pathway),ncol=1)
for (i in c(1:length(all_genes_in_pathway))){
  matlog2fc[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_pathway[i],'logFC'])
  matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_pathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_pathway,dataframe_expression_table[all_genes_in_pathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_M9"
rownames(matadjpval)<-paste(all_genes_in_pathway,dataframe_expression_table[all_genes_in_pathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_M9"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title="Hyphal Growth Genes")
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
pathway="hyphal_growth_genes"
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("NON_KEGG_Requested", pathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```

### __12. NON KEGG PATHWAY:  CWI Pathway Genes__

__Converted list of genes from Tamires:__

```{r cwi genes from Tamires m9}
custom_genes <- read.table("cwi_genes.txt",header=F)
custom_genes_info<-gene_info[which(gene_info$locus_tag %in% custom_genes[,1]),]
custom_genes_exp<-cbind(custom_genes_info,dataframe_expression_table[custom_genes_info$locus_tag,])
create_dt(custom_genes_exp)
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_pathway<-custom_genes[,1]
matlog2fc <- matrix(nrow=length(all_genes_in_pathway),ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_pathway),ncol=1)
for (i in c(1:length(all_genes_in_pathway))){
  matlog2fc[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_pathway[i],'logFC'])
  matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_pathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_pathway,dataframe_expression_table[all_genes_in_pathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_M9"
rownames(matadjpval)<-paste(all_genes_in_pathway,dataframe_expression_table[all_genes_in_pathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_M9"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title="CWI Pathway Genes")
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
pathway="cwi_pathway_genes"
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("NON_KEGG_Requested", pathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```


### __13. NON KEGG PATHWAY:  RAM Pathway Genes__

__Converted list of genes from Tamires:__

```{r ram genes from Tamires m9}
custom_genes <- read.table("ram_genes.txt",header=F)
custom_genes_info<-gene_info[which(gene_info$locus_tag %in% custom_genes[,1]),]
custom_genes_exp<-cbind(custom_genes_info,dataframe_expression_table[custom_genes_info$locus_tag,])
create_dt(custom_genes_exp)
```

__Heatmap of all genes in pathway__

Note that the plot has been saved as a PDF file and is not included in the report.

```{r,echo=FALSE}
all_genes_in_pathway<-custom_genes[,1]
matlog2fc <- matrix(nrow=length(all_genes_in_pathway),ncol = 1)
matadjpval<-matrix(nrow=length(all_genes_in_pathway),ncol=1)
for (i in c(1:length(all_genes_in_pathway))){
  matlog2fc[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_pathway[i],'logFC'])
  matadjpval[i,1]<-as.numeric(results_CaKp_vs_Ca_M9[all_genes_in_pathway[i],'adj.P.Val'])
}
rownames(matlog2fc)<-paste(all_genes_in_pathway,dataframe_expression_table[all_genes_in_pathway,"symbol"],sep="-")
colnames(matlog2fc)<-"CaKp_vs_Ca_M9"
rownames(matadjpval)<-paste(all_genes_in_pathway,dataframe_expression_table[all_genes_in_pathway,"symbol"],sep="-")
colnames(matadjpval)<-"CaKp_vs_Ca_M9"

# Create the heatmap
H <- Gene_Heatmap(matlog2fc,matadjpval,title="RAM Pathway Genes")
```

```{r, results='hide'}
# Save to PDF, Set height dynamically
pathway="ram_pathway_genes"
nr <- nrow(matlog2fc)
pdf(
  file.path(OUTPUT_DIR, paste("NON_KEGG_Requested", pathway, colnames(matlog2fc)[1], "heatmap.pdf", sep = "_")),
    width = 8,
    height = 0.1969 * nr + 3.3150
)
draw(H)
dev.off()
```



## __Klebsiella pneumoniae genes__
```{r subsetting kp}
txi_kp<-list()
txi_kp$abundance<-txi$abundance[grep("KPN",rownames(txi$abundance)),grep("K",colnames(txi$abundance))]
txi_kp$counts<-txi$counts[grep("KPN",rownames(txi$counts)),grep("K",colnames(txi$counts))]
txi_kp$length<-txi$length[grep("KPN",rownames(txi$length)),grep("K",colnames(txi$length))]
txi_kp$countsFromAbundance<-txi$countsFromAbundance
```

  
### __Total reads per library__
  
  CaKp mix culture libraries have much lower total number of reads assigned to Klebsiella genes compared to Kp single culture libraries.

```{r kp total reads lib}
barplot_data<-data.frame(
  counts=colSums(txi_kp$counts),
  species=metadata$group_name_2[grep("K",metadata$species_media)],
  media=metadata$group_name_1[grep("K",metadata$species_media)],
  names=metadata$sample_name_in_report[grep("K",metadata$species_media)]
)
p <- ggplot(data=barplot_data,aes(x=names,y=counts,color=media,fill=species)) + 
  geom_bar(stat="identity") + 
  scale_fill_brewer(palette="Set1") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_color_brewer(palette="Dark2")
p
```
__Create a DGEList__

```{r kp create DGEList}
suppressPackageStartupMessages(library(edgeR))
y <- DGEList(txi_kp$counts)
```

### __Removing genes that are not expressed__

__Specifying the factors in the experimental design__

```{r kp factors, echo=FALSE}
SPECIES_MEDIA<-metadata$species_media[grep("K",metadata$species_media)]
```

```{r kp design initial}
design<-model.matrix(~0 + SPECIES_MEDIA)
```

__Flter low expressed genes using edgeR's filterByExpr__
```{r kp exressionpfiltering}

keep<-filterByExpr(y,design = design)
filtered_y<-y[keep,]

```

No. of genes before filtering: __`r nrow(y)`__

No. of genes after filtering out low expression genes: __`r nrow(filtered_y)`__

```{r kp plotexpfilter, echo=FALSE}
legend_names=colnames(y)
#plot log2 CPM for genes before and after filtering out low expression genes
#before filtering
lcpm=cpm(y$counts,log = TRUE)
L <- mean(y$samples$lib.size) * 1e-6
M <- median(y$samples$lib.size) * 1e-6
lcpm.cutoff <- log2(10/M + 2/L)
nsamples <- ncol(y)
col <- hcl.colors(nsamples, palette = "zissou1", alpha = NULL, rev = FALSE, fixup = TRUE)
par(mfrow=c(1,2))
plot(density(lcpm[,1]), col=col[1], lwd=2, ylim=c(0,0.26), las=2, main="", xlab="",cex.axis=0.8,cex.lab=0.8)
title(main="A. Raw data", xlab="Log-cpm",cex.main=0.8,cex.lab=0.8)
abline(v=lcpm.cutoff, lty=3)
for (i in 2:nsamples){
  den <- density(lcpm[,i])
  lines(den$x, den$y, col=col[i], lwd=2)
}
legend("topright", legend_names, text.col=col, bty="n",cex=0.4)
#after filtering
lcpm <- cpm(filtered_y, log=TRUE)
plot(density(lcpm[,1]), col=col[1], lwd=2, ylim=c(0,0.26), las=2, main="", xlab="",cex.axis=0.8,cex.lab=0.8)
title(main="B. Filtered data", xlab="Log-cpm",cex.main=0.8,cex.lab=0.8)
abline(v=lcpm.cutoff, lty=3)
for (i in 2:nsamples){
  den <- density(lcpm[,i])
  lines(den$x, den$y, col=col[i], lwd=2)
}
legend("topright", legend_names, text.col=col, bty="n",cex=0.4)
```

### __Expression data__

```{r kp additional gene info}
#genes info lookup
feature_table_file <- file.path(ANNOTATION_DIR, "GCF_000016305.1_ASM1630v1_feature_table.txt") # created from cds header lines
gene_feature_table<- read.table(feature_table_file,
                             header = TRUE,
                             sep="\t",
                             stringsAsFactors = FALSE,
                             comment.char ="",
                             quote=""
                          )
gene_info<-gene_feature_table[which(gene_feature_table[,1]=="CDS"),c("locus_tag","symbol","name")]
```

Counts per million expression values are included in the table below. 
```{r kp expdata, echo=FALSE}
#create downloadable normalized counts data table
expression_table<-cpm(filtered_y,log=TRUE) #log cpm
dataframe_expression_table_exp<-as.data.frame(expression_table)
rownames(gene_info)<-gene_info$locus_tag
dataframe_expression_table<-merge(gene_info,dataframe_expression_table_exp,by = c(0,0),all.y=TRUE)
rownames(dataframe_expression_table)<-dataframe_expression_table$Row.names

#create_dt(dataframe_expression_table[,-which(names(dataframe_expression_table)=='tx_ids')])
## change values to non log2
dataframe_expression_table[,grep("^C",colnames(dataframe_expression_table))]<-2^dataframe_expression_table[,grep("^K",colnames(dataframe_expression_table))]
kp_counts_per_million_expression_table<-dataframe_expression_table[,-c(1)]
create_dt(kp_counts_per_million_expression_table)
save_xlsx(kp_counts_per_million_expression_table)
```



## __KP genes: Exploratory data analysis__

### __Top 1000 most variable genes for the exploratory data analysis__

```{r kp cov}
cov_lcpm<-apply(lcpm,1,function(x){sd(x)/mean(x)})
num_highest_cov=1000
keep_2<-names(head(sort(cov_lcpm,decreasing=TRUE),num_highest_cov))
top_lcpm<-lcpm[keep_2,]
kp_top_lcpm<-top_lcpm
```

#### __Heatmap with hierarchical clustering between samples (across all genes)__

`r nrow(top_lcpm)` genes with the highest coefficients of variation were used for the analyses in this section

Euclidean distances between samples were calculated with log2 counts per million (CPM) values across all genes
```{r kp similarityclus, echo=FALSE,fig.height=6,fig.width=9}
library(pheatmap)
library(RColorBrewer)
colors <- colorRampPalette(rev(brewer.pal(9, "Blues")))(255)
fs=9
lcpm_dist<-dist(t(top_lcpm))
lcpm_distMatrix<-as.matrix(lcpm_dist)


heatmapnames <- metadata$sample_name_in_report[grep("K",metadata$species_media)]

# define the annotation
annotation_row = data.frame(
                          culture = factor(metadata$group_name_2[grep("K",metadata$species_media)]),
                          media = factor(metadata$group_name_1[grep("K",metadata$species_media)])
                        )
rownames(annotation_row) = heatmapnames
annotation_colour<-list()
colour_plot<-hcl.colors(5, palette = "zissou1", alpha = NULL, rev = TRUE, fixup = TRUE)
j=1
for (i in c(1:ncol(annotation_row))){
  n=length(levels(annotation_row[,names(annotation_row)[i]]))
  annot_col=colour_plot[j:(j+n-1)]
  j=j+n
  names(annot_col)<-levels(annotation_row[,names(annotation_row)[i]])
  annotation_colour[[i]]<-annot_col  
}
names(annotation_colour)<-names(annotation_row)

rownames(lcpm_distMatrix)<-heatmapnames
colnames(lcpm_distMatrix)<-heatmapnames
#par(mar=c(1,1,1,1))
pheatmap(lcpm_distMatrix,
         clustering_distance_rows = lcpm_dist,
         clustering_distance_cols = lcpm_dist,
         col = colors,
         fontsize_row = fs, fontsize_col = fs,
         cellheight = 15, cellwidth = 15,
         annotation_row = annotation_row,
         annotation_colors = annotation_colour,
         angle_col=90
         )
```

#### __Principal Component Analysis (PCA)__

```{r kp PCA generation, message=FALSE, eval=TRUE, include=TRUE, echo=FALSE}
additional_col <- data.frame(culture = factor(metadata$group_name_2[grep("K",metadata$species_media)]),media = factor(metadata$group_name_1[grep("K",metadata$species_media)]), stringsAsFactors =FALSE)
data_for_pca <- cbind(additional_col,t(as.matrix(top_lcpm)))
pca_object <- prcomp(t(as.matrix(top_lcpm)), center = TRUE, scale=TRUE)
summ<-summary(pca_object)
```

About `r round(summ$importance[3,2]*100,1)`% of the total variation is explained in PC1 and PC2 of the PCA.


```{r kp pca plot, echo=FALSE,fig.height=6,fig.width=9}
par(mar=c(3,3,3,3))
p<-autoplot(pca_object,data=data_for_pca,colour='culture',shape='media',label=TRUE,size=3,label.size=4)
p+scale_color_manual(values=annotation_colour$culture)
```


## __KP genes: Differential Expression Analysis__

__Comparisons of interest__

The following comparisons are of interest for Klebsiella

b) Cakp vs Kp (from RPMI group)
d) Candida_Kp vs Klebsiella (from M9 group)


__Normalization__

[TMM normalization](https://doi.org/10.1186/gb-2010-11-3-r25 "Robinson, M.D., Oshlack, A. A scaling normalization method for differential expression analysis of RNA-seq data. Genome Biol 11, R25 (2010).") for between sample comparisons  

```{r kp normalization}

# normalize 
dge <- calcNormFactors(filtered_y)
lcpm=cpm(dge$counts,log = TRUE)
boxplot(lcpm, las=2,ylab="log2CountsPerMillion",cex.axis=0.45,ylim=c(-10,20),cex.main=0.8,cex.lab=0.8,pch='.')
```

#### __Heatmap of gene expression for RPMI samples (across all genes)__

```{r kp gene expression heatmap RPMI}
lcpm_heatmap<-lcpm[,which(colnames(lcpm) %in% metadata[which(metadata[,"group_name_1"]=="RPMI"),]$sample_name_in_report)]
lcpm_heatmap<-lcpm_heatmap[apply(lcpm_heatmap,1,sd)>0,] #drop genes invariant across this sample subset (cannot be row-scaled)
p<-pheatmap(lcpm_heatmap, scale = "row",show_rownames = FALSE,heatmap_legend_param = list(title = "Row Z-score"),name="RPMI",col=rev(brewer.pal(n = 8, name = "RdBu")),cellwidth=15)
pdf(file.path(OUTPUT_DIR,"Kp_gene_expression_RPMI_heatmap.pdf"),height=9,width=8)
p
dev.off()
p
```

#### __Heatmap of gene expression for M9 samples (across all genes)__

```{r kp gene expression heatmap M9}
lcpm_heatmap<-lcpm[,which(colnames(lcpm) %in%  gsub(" ", "",metadata[which(metadata[,"group_name_1"]=="M9"),]$sample_name_in_report))]
lcpm_heatmap<-lcpm_heatmap[apply(lcpm_heatmap,1,sd)>0,] #drop genes invariant across this sample subset (cannot be row-scaled)
p<-pheatmap(lcpm_heatmap, scale = "row",show_rownames = FALSE,heatmap_legend_param = list(title = "Row Z-score"),name="M9",col=rev(brewer.pal(n = 8, name = "RdBu")),cellwidth=15)
pdf(file.path(OUTPUT_DIR,"Kp_gene_expression_M9_heatmap.pdf"),height=9,width=8)
p
dev.off()
p
```

__Thresholds__

```{r kp thresholds}
#DEG Thresholds
adj_p_value_thresh=0.05
FC=1.5
FC_thresh=log2(FC)
```

__Transform__

[limma-voom](https://doi.org/10.1093/nar/gkv007 "Ritchie ME, Phipson B, Wu D, Hu Y, Law CW, Shi W, Smyth GK (2015). “limma powers differential expression analyses for RNA-sequencing and microarray studies.” Nucleic Acids Research, 43(7), e47.") transformation to remove heteroscedascity

```{r kp design modif, echo=FALSE}
colnames(design)<-gsub("SPECIES_MEDIA","",colnames(design))
colnames(design)<-gsub(" ","_",colnames(design))
colnames(design)<-gsub("_mix/","",colnames(design))

```

```{r kp voom} 
par(mfrow=c(1,2))
v <- voom(dge, design,plot=T)
fit<-lmFit(v,design)
efit <- eBayes(fit)
plotSA(efit, main="Final model: Mean-variance trend")
```

### __Comparison: A. Cakp vs Kp (from RPMI group)__

#### __Contrast matrix__
```{r kp Contrast cakp vs kp rpmi}
contrast.matrix <- makeContrasts(CaKp_vs_Kp_RPMI = CaKp_mix__RPMI - Kp_single__RPMI, levels=colnames(design))
```

#### __Significant DEGs__
```{r kp DEG cakp vs kp rpmi}
fit_2 <- contrasts.fit(fit,contrast.matrix)
fit_2 <- eBayes(fit_2)

##Benjamini-Hochberg method used to adjust the p-values for multiple testing
results<-topTable(fit_2, adjust="BH",number=nrow(filtered_y))
results_CaKp_vs_Kp_RPMI<-results
#all significant DEGs based on p-value and FC thresholds
results_CaKp_vs_Kp_RPMI_DEGS<-results[which((abs(results$logFC)>FC_thresh&results$adj.P.Val<adj_p_value_thresh)),]
results_CaKp_vs_Kp_RPMI_UPREG<-results[which((results$logFC>FC_thresh&results$adj.P.Val<adj_p_value_thresh)),]
results_CaKp_vs_Kp_RPMI_DOWNREG<-results[which((results$logFC<(-FC_thresh)&results$adj.P.Val<adj_p_value_thresh)),]
CaKp_vs_Kp_RPMI_DEGS=rownames(results_CaKp_vs_Kp_RPMI_DEGS)
CaKp_vs_Kp_RPMI_DEGS_UPREG=rownames(results_CaKp_vs_Kp_RPMI_UPREG)
CaKp_vs_Kp_RPMI_DEGS_DOWNREG=rownames(results_CaKp_vs_Kp_RPMI_DOWNREG)
add_attributes<-c("symbol","name",colnames(dataframe_expression_table)[grep("C",colnames(dataframe_expression_table))])
results_CaKp_vs_Kp_RPMI_DEGS[,add_attributes]<-dataframe_expression_table[match(CaKp_vs_Kp_RPMI_DEGS,dataframe_expression_table[,"locus_tag"]),add_attributes]
create_dt(results_CaKp_vs_Kp_RPMI_DEGS)
save_xlsx(results_CaKp_vs_Kp_RPMI_DEGS)
```

__Summary table of Up/Down/NotSig__
```{r kp summary degs cakp vs kp rpmi,echo=FALSE}
summary_table<-summary(decideTests(fit_2,lfc=FC_thresh,p.value=adj_p_value_thresh))
colnames(summary_table)<-gsub("=.*","",colnames(summary_table))
summary_table
```
__Volcano plots__
```{r kp volcanos cakp vs kp rpmi,fig.height=7,fig.width=12, out.width = "\\textwidth*10", out.height = "\\textheight*800"}
res<-results
res$symbol<-dataframe_expression_table[rownames(results),"symbol"]
res_sig<-res[res$adj.P.Val<adj_p_value_thresh,]
sig_down<-res_sig[which(res_sig$logFC<(-FC_thresh)),]
top_sig_down<-head(sig_down$symbol[sig_down$symbol!=""],top_n)
fc_down<-sig_down[order(sig_down$logFC),]
top_fc_down<-head(fc_down$symbol[fc_down$symbol!=""],top_n)
sig_up<-res_sig[which(res$logFC>(FC_thresh)),]
top_sig_up<-head(sig_up$symbol[sig_up$symbol!=""],top_n)
fc_up<-sig_up[order(sig_up$logFC,decreasing=TRUE),]
top_fc_up<-head(fc_up$symbol[fc_up$symbol!=""],top_n)
genes_of_interest<-c(top_sig_down,top_sig_up,top_fc_down,top_fc_up)
p<-plotVolcano(results_obj = res,genes_of_interest = genes_of_interest,plot_title=colnames(contrast.matrix),FC_thresh=log2(FC),adj_p_value_thresh=adj_p_value_thresh)
vplots<-list()
vplots[[length(vplots)+1]]<-p
pdf(file.path(OUTPUT_DIR,"CaKp_vs_Kp_RPMI_volcanoplot.pdf"),height=8,width=8)
p
dev.off()
p

```

#### __Heatmap of gene expression for RPMI samples (Top Upregulated and Downregulated DEGs)__

```{r kp top deg heatmap RPMI}
top_n=50
top_fc_down<-head(fc_down,top_n)
top_fc_up<-head(fc_up,top_n)
top_fc_genes<-c(rownames(top_fc_down),rownames(top_fc_up))
lcpm_heatmap<-lcpm[top_fc_genes,which(colnames(lcpm) %in% metadata[which(metadata[,"group_name_1"]=="RPMI"),]$sample_name_in_report)]
rownames(lcpm_heatmap)<-c(paste(rownames(top_fc_down),top_fc_down$symbol,sep="-"),paste(rownames(top_fc_up),top_fc_up$symbol,sep="-"))
p<-pheatmap(lcpm_heatmap, scale = "row",show_rownames = TRUE,heatmap_legend_param = list(title = "Row Z-score"),name="RPMI",col=rev(brewer.pal(n = 8, name = "RdBu")),cellwidth=8,cellheight=8, fontsize=8, fontsize_row=6,cluster_cols = FALSE)
pdf(file.path(OUTPUT_DIR,"Kp_top_DEGs_expression_RPMI_heatmap.pdf"),height=13,width=5)
p
dev.off()
p
```

### __Comparison: C. Cakp vs Kp (from M9 group)__

#### __Contrast matrix__
```{r kp Contrast cakp vs kp m9}
contrast.matrix <- makeContrasts(CaKp_vs_Kp_M9 = CaKp_mix__M9 - Kp_single__M9, levels=colnames(design))
```

#### __Significant DEGs__
```{r kp DEG cakp vs kp m9}
fit_2 <- contrasts.fit(fit,contrast.matrix)
fit_2 <- eBayes(fit_2)

##Benjamini-Hochberg method used to adjust the p-values for multiple testing
results<-topTable(fit_2, adjust="BH",number=nrow(filtered_y))
results_CaKp_vs_Kp_M9<-results
#all significant DEGs based on p-value and FC thresholds
results_CaKp_vs_Kp_M9_DEGS<-results[which((abs(results$logFC)>FC_thresh&results$adj.P.Val<adj_p_value_thresh)),]
results_CaKp_vs_Kp_M9_UPREG<-results[which((results$logFC>FC_thresh&results$adj.P.Val<adj_p_value_thresh)),]
results_CaKp_vs_Kp_M9_DOWNREG<-results[which((results$logFC<(-FC_thresh)&results$adj.P.Val<adj_p_value_thresh)),]
CaKp_vs_Kp_M9_DEGS=rownames(results_CaKp_vs_Kp_M9_DEGS)
CaKp_vs_Kp_M9_DEGS_UPREG=rownames(results_CaKp_vs_Kp_M9_UPREG)
CaKp_vs_Kp_M9_DEGS_DOWNREG=rownames(results_CaKp_vs_Kp_M9_DOWNREG)
add_attributes<-c("symbol","name",colnames(dataframe_expression_table)[grep("C",colnames(dataframe_expression_table))])
results_CaKp_vs_Kp_M9_DEGS[,add_attributes]<-dataframe_expression_table[match(CaKp_vs_Kp_M9_DEGS,dataframe_expression_table[,"locus_tag"]),add_attributes]
create_dt(results_CaKp_vs_Kp_M9_DEGS)
save_xlsx(results_CaKp_vs_Kp_M9_DEGS)
```

__Summary table of Up/Down/NotSig__
```{r kp summary degs tables cakp vs kp m9,echo=FALSE}
summary_table<-summary(decideTests(fit_2,lfc=FC_thresh,p.value=adj_p_value_thresh))
colnames(summary_table)<-gsub("=.*","",colnames(summary_table))
summary_table
```
__Volcano plots__
```{r kp volcanos cakp vs kp m9,fig.height=7,fig.width=12, out.width = "\\textwidth*10", out.height = "\\textheight*800"}
res<-results
res$symbol<-dataframe_expression_table[rownames(results),"symbol"]
res_sig<-res[res$adj.P.Val<adj_p_value_thresh,]
sig_down<-res_sig[which(res_sig$logFC<(-FC_thresh)),]
top_sig_down<-head(sig_down$symbol[sig_down$symbol!=""],top_n)
fc_down<-sig_down[order(sig_down$logFC),]
top_fc_down<-head(fc_down$symbol[fc_down$symbol!=""],top_n)
sig_up<-res_sig[which(res$logFC>(FC_thresh)),]
top_sig_up<-head(sig_up$symbol[sig_up$symbol!=""],top_n)
fc_up<-sig_up[order(sig_up$logFC,decreasing=TRUE),]
top_fc_up<-head(fc_up$symbol[fc_up$symbol!=""],top_n)
genes_of_interest<-c(top_sig_down,top_sig_up,top_fc_down,top_fc_up)
p<-plotVolcano(results_obj = res,genes_of_interest = genes_of_interest,plot_title=colnames(contrast.matrix),FC_thresh=log2(FC),adj_p_value_thresh=adj_p_value_thresh)
vplots<-list()
vplots[[length(vplots)+1]]<-p
pdf(file.path(OUTPUT_DIR,"CaKp_vs_Kp_M9_volcanoplot.pdf"),height=8,width=8)
p
dev.off()
p

```

#### __Heatmap of gene expression for M9 samples (Top Upregulated and Downregulated DEGs)__

```{r kp top deg heatmap M9}
top_n=50
top_fc_down<-head(fc_down,top_n)
top_fc_up<-head(fc_up,top_n)
top_fc_genes<-c(rownames(top_fc_down),rownames(top_fc_up))
lcpm_heatmap<-lcpm[top_fc_genes,which(colnames(lcpm) %in%  gsub(" ", "",metadata[which(metadata[,"group_name_1"]=="M9"),]$sample_name_in_report))]
rownames(lcpm_heatmap)<-c(paste(rownames(top_fc_down),top_fc_down$symbol,sep="-"),paste(rownames(top_fc_up),top_fc_up$symbol,sep="-"))
p<-pheatmap(lcpm_heatmap, scale = "row",show_rownames = TRUE,heatmap_legend_param = list(title = "Row Z-score"),name="M9",col=rev(brewer.pal(n = 8, name = "RdBu")),cellwidth=8,cellheight=8, fontsize=8, fontsize_row=6,cluster_cols = FALSE)
pdf(file.path(OUTPUT_DIR,"Kp_top_DEGs_expression_M9_heatmap.pdf"),height=13,width=5)
p
dev.off()
p
```


#### __Cakp vs Kp DEGs venn__

```{r overlap comp cakp vs kp ,echo=FALSE,fig.height=7,fig.width=12,out.width='85%'}
signatures<-list(CaKp_vs_Kp_RPMI_DEGS_UPREG=CaKp_vs_Kp_RPMI_DEGS_UPREG,CaKp_vs_Kp_RPMI_DEGS_DOWNREG=CaKp_vs_Kp_RPMI_DEGS_DOWNREG,CaKp_vs_Kp_M9_DEGS_UPREG=CaKp_vs_Kp_M9_DEGS_UPREG,CaKp_vs_Kp_M9_DEGS_DOWNREG=CaKp_vs_Kp_M9_DEGS_DOWNREG)

library(eulerr)
vennfit <- euler(signatures)
plot(vennfit,
     fills = c(brewer.pal(n = 8, name = "RdBu")[2],brewer.pal(n = 8, name = "RdBu")[7],brewer.pal(n = 8, name = "RdBu")[3], brewer.pal(n = 8, name = "RdBu")[6]),
     edges = TRUE,
     fontsize = 8,
     quantities = list(fontsize = 8))
```

```{r kp gene lists}
write_venn_genelists(signatures = signatures,sig_names_interest = names(signatures), outbasedir=file.path(OUTPUT_DIR,"CaKp_vs_Kp__venndiagram_genes"))
```

### __GO Enrichment analysis with TopGO__ 

In this section the R package   [TopGO](https://bioconductor.org/packages/release/bioc/html/topGO.html "Alexa A, Rahnenfuhrer J (2022). topGO: Enrichment Analysis for Gene Ontology. R package version 2.49.0.") was used to reanalyse GO Biological Process terms, accounting for the relationship between parent and child GO terms. The algorithm used was `weight01`. Statistical significance was tested with the Fisher's exact test and multiple testing correction was performed using the Benjamini-Hochberg method.


```{r kp topGO setup}
#GO term look up
gomappingsfile="GCF_000016305.1_ASM1630v1_genomic_geneid2go_mappings.tsv"
geneID2GO<-readMappings(file=file.path(ANNOTATION_DIR, gomappingsfile))
gotable<-read.table(file.path(ANNOTATION_DIR, gomappingsfile))
colnames(gotable)<-c("locus_tag","go_terms")
create_dt(gotable)
```

#### __RPMI__

```{r kp topGO degs RPMI}
media="RPMI"
#1. Run TopGO for each signature
geneNames<-names(geneID2GO)
GOdata_list<-lapply(grep(media,names(signatures),value = T),function(comp){
  interestingGenes<-signatures[[comp]]
  geneList<-factor(as.integer(geneNames %in% interestingGenes))
  names(geneList)<-geneNames
  GOdata <- new('topGOdata',
                ontology = 'BP',
                allGenes = geneList,
                annot = annFUN.gene2GO,
                gene2GO = geneID2GO,
                nodeSize = 10)
  return(GOdata)  
})
names(GOdata_list)<-grep(media,names(signatures),value = T)

```

```{r kp significant terms rpmi}

#2. Apply Fisher exact test with weight01 algorithm from TopGO
test.stat <- new("weight01Count", testStatistic = GOFisherTest, name = "Fisher test")

#3.-7. Run topGO_wrapper that produces a list of tables for each comparison for all significant terms and a comparison plot 
topGO_results<-topGO_wrapper(GOdata_list,test.stat,FDR_thres=0.05)

# Table of genes in each GO term
topGO_results<-genes_per_pathway(GOdata_list,signatures,topGO_results=topGO_results)

##Write file as excel file with multiple worksheets
openxlsx::write.xlsx(topGO_results$result_list,file.path(OUTPUT_DIR,paste(media,"kp_degs_topGO_results.xlsx",sep="_")),colNames=TRUE,rowNames=FALSE)
```


```{r kp topgo plot rpmi,echo=FALSE,fig.width=8,fig.height=6+length(unique(topGO_results$sigplot$data$Term))/4}
topGO_results$sigplot
```

#### __M9__

```{r kp topGO degs M9}
media="M9"
#1. Run TopGO for each signature
geneNames<-names(geneID2GO)
GOdata_list<-lapply(grep(media,names(signatures),value = T),function(comp){
  interestingGenes<-signatures[[comp]]
  geneList<-factor(as.integer(geneNames %in% interestingGenes))
  names(geneList)<-geneNames
  GOdata <- new('topGOdata',
                ontology = 'BP',
                allGenes = geneList,
                annot = annFUN.gene2GO,
                gene2GO = geneID2GO,
                nodeSize = 10)
  return(GOdata)  
})
names(GOdata_list)<-grep(media,names(signatures),value = T)

```

```{r kp significant terms m9}
#2. Apply Fisher exact test with weight01 algorithm from TopGO
test.stat <- new("weight01Count", testStatistic = GOFisherTest, name = "Fisher test")

#3.-7. Run topGO_wrapper that produces a list of tables for each comparison for all significant terms and a comparison plot 
topGO_results<-topGO_wrapper(GOdata_list[grep(media,names(GOdata_list))],test.stat,FDR_thres=0.05)

# Table of genes in each GO term
topGO_results<-genes_per_pathway(GOdata_list[grep(media,names(GOdata_list))],signatures[grep(media,names(signatures))],topGO_results=topGO_results)

##Write file as excel file with multiple worksheets
openxlsx::write.xlsx(topGO_results$result_list,file.path(OUTPUT_DIR,paste(media,"kp_degs_topGO_results.xlsx",sep="_")),colNames=TRUE,rowNames=FALSE)
```

```{r kp topgo plot media,echo=FALSE,fig.width=8,fig.height=6+length(unique(topGO_results$sigplot$data$Term))/4}
topGO_results$sigplot
```


### __KEGG Pathways analysis: Over Representation/Enrichment analysis__

See the KEGG reproducibility note in the *C. albicans* over-representation section above.

#### __Old locus tag__

```{r table locus tag}
info_locus_tag<-cbind(gene_feature_table$locus_tag,gene_feature_table$attributes)
info_locus_tag<-info_locus_tag[grep("old_locus_tag",info_locus_tag[,2]),]
info_locus_tag[,2]<-gsub('old_locus_tag=','',gsub('pseudo;', '', info_locus_tag[,2]))
colnames(info_locus_tag)<-c('locus_tag','old_locus_tag')
create_dt(info_locus_tag)
rownames(info_locus_tag)<-info_locus_tag[,'locus_tag']
```

#### __All DEGs RPMI__

```{r convert ids to old ora kp rpmi}
geneset=rownames(results_CaKp_vs_Kp_RPMI_DEGS)
old_ids_genesset<-info_locus_tag[geneset[which(geneset %in% rownames(info_locus_tag))],"old_locus_tag"]
```

`r length(old_ids_genesset)` out of `r length(geneset)` genes could be converted to the old locus tag of KPN.

```{r ora kp rpmi}
keggEnr<-enrichKEGG(old_ids_genesset,organism='kpn',pvalueCutoff = 0.05)
create_dt(keggEnr@result)
```
```{r ora kp rpmi plot,fig.width=10,fig.height=(4+dim(keggEnr)[1]/2)}
dotplot_scale(keggEnr, title="All DEGs RPMI",font.size=10)
```


#### __Upregulated DEGs RPMI__


```{r convert ids to old ora kp rpmi up}
geneset=rownames(results_CaKp_vs_Kp_RPMI_UPREG)
old_ids_genesset<-info_locus_tag[geneset[which(geneset %in% rownames(info_locus_tag))],"old_locus_tag"]
```

`r length(old_ids_genesset)` out of `r length(geneset)` genes could be converted to the old locus tag of KPN.


```{r ora kp rpmi up}
keggEnr<-enrichKEGG(old_ids_genesset,organism='kpn',pvalueCutoff = 0.05)
create_dt(keggEnr@result)
```

```{r ora kp rpmi up plot,fig.width=10,fig.height=(4+dim(keggEnr)[1]/2)}
dotplot_scale(keggEnr, title="Upreg DEGs RPMI",font.size=10)
```

#### __Downregulated DEGs RPMI__


```{r convert ids to old ora kp rpmi down}
geneset=rownames(results_CaKp_vs_Kp_RPMI_DOWNREG)
old_ids_genesset<-info_locus_tag[geneset[which(geneset %in% rownames(info_locus_tag))],"old_locus_tag"]
```

`r length(old_ids_genesset)` out of `r length(geneset)` genes could be converted to the old locus tag of KPN.


```{r ora kp rpmi down}
keggEnr<-enrichKEGG(old_ids_genesset,organism='kpn',pvalueCutoff = 0.05)
create_dt(keggEnr@result)
```

```{r ora kp rpmi down plot,fig.width=10,fig.height=(4+dim(keggEnr)[1]/2)}
dotplot_scale(keggEnr, title="Downreg DEGs RPMI",font.size=10)
```


#### __All DEGs M9__

```{r convert ids to old ora kp M9}
geneset=rownames(results_CaKp_vs_Kp_M9_DEGS)
old_ids_genesset<-info_locus_tag[geneset[which(geneset %in% rownames(info_locus_tag))],"old_locus_tag"]
```

`r length(old_ids_genesset)` out of `r length(geneset)` genes could be converted to the old locus tag of KPN.

```{r ora kp M9}
keggEnr<-enrichKEGG(old_ids_genesset,organism='kpn',pvalueCutoff = 0.05)
create_dt(keggEnr@result)
```
```{r ora kp M9 plot,fig.width=10,fig.height=(4+dim(keggEnr)[1]/2)}
dotplot_scale(keggEnr, title="All DEGs M9",font.size=10)
```


#### __Upregulated DEGs M9__


```{r convert ids to old ora kp M9 up}
geneset=rownames(results_CaKp_vs_Kp_M9_UPREG)
old_ids_genesset<-info_locus_tag[geneset[which(geneset %in% rownames(info_locus_tag))],"old_locus_tag"]
```

`r length(old_ids_genesset)` out of `r length(geneset)` genes could be converted to the old locus tag of KPN.


```{r ora kp M9 up}
keggEnr<-enrichKEGG(old_ids_genesset,organism='kpn',pvalueCutoff = 0.05)
create_dt(keggEnr@result)
```

```{r ora kp M9 up plot,fig.width=10,fig.height=(4+dim(keggEnr)[1]/2)}
dotplot_scale(keggEnr, title="Upreg DEGs M9",font.size=10)
```

#### __Downregulated DEGs M9__


```{r convert ids to old ora kp M9 down}
geneset=rownames(results_CaKp_vs_Kp_M9_DOWNREG)
old_ids_genesset<-info_locus_tag[geneset[which(geneset %in% rownames(info_locus_tag))],"old_locus_tag"]
```

`r length(old_ids_genesset)` out of `r length(geneset)` genes could be converted to the old locus tag of KPN.


```{r ora kp M9 down}
keggEnr<-enrichKEGG(old_ids_genesset,organism='kpn',pvalueCutoff = 0.05)
create_dt(keggEnr@result)
```

```{r ora kp M9 down plot,fig.width=10,fig.height=(4+dim(keggEnr)[1]/2)}
dotplot_scale(keggEnr, title="Downreg DEGs M9",font.size=10)
```

### __KEGG Pathways analysis: Gene Set Enrichment Analysis (GSEA)__

See the KEGG reproducibility note in the *C. albicans* over-representation section above.

Note only the top 10 pathways are shown in the figures.

#### __RPMI__ 

```{r convert ids to old ora kp RPMI gsea}
geneList<-as.matrix(results_CaKp_vs_Kp_RPMI$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Kp_RPMI)
geneset=rownames(results_CaKp_vs_Kp_RPMI)
old_ids_genesset<-info_locus_tag[geneset[which(geneset %in% rownames(info_locus_tag))],"old_locus_tag"]
geneList<-as.matrix(geneList[geneset[which(geneset %in% rownames(info_locus_tag))],])
rownames(geneList)<-old_ids_genesset
```

`r length(old_ids_genesset)` out of `r length(geneset)` genes could be converted to the old locus tag of KPN.


```{r gsea kp rpmi,fig.height=10,fig.width=20.5}
keggGs<-gseKEGG(geneList = sort(geneList[,1],decreasing = T), organism = 'kpn', minGSSize = 10, pvalueCutoff = 0.05, verbose = F,eps =0)
create_dt(keggGs@result)
```

```{r gsea kp rpmi plot hide, fig.height=10,fig.width=10,results='hide',echo=FALSE,fig.show='hide'}
final_plot<-modgseaplot(keggGs,geneSetID,colourlist)
```

```{r gsea kp rpmi plot, fig.height=4,fig.width=13,echo=FALSE,out.width = "200%"}
print(final_plot)
```




#### __M9__ 

```{r convert ids to old ora kp M9 gsea}
geneList<-as.matrix(results_CaKp_vs_Kp_M9$logFC)
rownames(geneList)<-rownames(results_CaKp_vs_Kp_M9)
geneset=rownames(results_CaKp_vs_Kp_M9)
old_ids_genesset<-info_locus_tag[geneset[which(geneset %in% rownames(info_locus_tag))],"old_locus_tag"]
geneList<-as.matrix(geneList[geneset[which(geneset %in% rownames(info_locus_tag))],])
rownames(geneList)<-old_ids_genesset
```

`r length(old_ids_genesset)` out of `r length(geneset)` genes could be converted to the old locus tag of KPN.


```{r gsea kp M9,fig.height=10,fig.width=20.5}
keggGs<-gseKEGG(geneList = sort(geneList[,1],decreasing = T), organism = 'kpn', minGSSize = 10, pvalueCutoff = 0.05, verbose = F,eps =0)
create_dt(keggGs@result)
```


```{r gsea kp m9 plot hide, fig.height=10,fig.width=10,results='hide',echo=FALSE,fig.show='hide'}
final_plot<-modgseaplot(keggGs,geneSetID,colourlist)
```

```{r gsea kp m9 plot, fig.height=4,fig.width=14,echo=FALSE,out.width = "200%"}
print(final_plot)
```

## __Combined data__

### __Master table of expression and DE comparisons__

#### __Combining Ca DE analysis results with Ca genes' expression__
```{r ca master}
ca_counts_per_million_expression_table_with_de_comp<-ca_counts_per_million_expression_table
ca_counts_per_million_expression_table_with_de_comp$DE__CaKp_vs_Ca_RPMI_logFC=results_CaKp_vs_Ca_RPMI[rownames(ca_counts_per_million_expression_table),"logFC"]
ca_counts_per_million_expression_table_with_de_comp$DE__CaKp_vs_Ca_RPMI_adj.P.Val=results_CaKp_vs_Ca_RPMI[rownames(ca_counts_per_million_expression_table),"adj.P.Val"]
ca_counts_per_million_expression_table_with_de_comp$DE__CaKp_vs_Ca_M9_logFC=results_CaKp_vs_Ca_M9[rownames(ca_counts_per_million_expression_table),"logFC"]
ca_counts_per_million_expression_table_with_de_comp$DE__CaKp_vs_Ca_M9_adj.P.Val=results_CaKp_vs_Ca_M9[rownames(ca_counts_per_million_expression_table),"adj.P.Val"]
```

#### __Adding information on Ca transcription factors to the table__
```{r ca tf full list}
tf_list<-read.table(TFs_list_file)
ca_counts_per_million_expression_table_with_de_comp$Ca_TF<-rownames(ca_counts_per_million_expression_table) %in% tf_list[,1]
```

#### __Combining Kp DE analysis results with Kp genes' expression__
```{r kp master}
kp_counts_per_million_expression_table_with_de_comp<-kp_counts_per_million_expression_table
kp_counts_per_million_expression_table_with_de_comp$DE__CaKp_vs_Kp_RPMI_logFC=results_CaKp_vs_Kp_RPMI[rownames(kp_counts_per_million_expression_table),"logFC"]
kp_counts_per_million_expression_table_with_de_comp$DE__CaKp_vs_Kp_RPMI_adj.P.Val=results_CaKp_vs_Kp_RPMI[rownames(kp_counts_per_million_expression_table),"adj.P.Val"]
kp_counts_per_million_expression_table_with_de_comp$DE__CaKp_vs_Kp_M9_logFC=results_CaKp_vs_Kp_M9[rownames(kp_counts_per_million_expression_table),"logFC"]
kp_counts_per_million_expression_table_with_de_comp$DE__CaKp_vs_Kp_M9_adj.P.Val=results_CaKp_vs_Kp_M9[rownames(kp_counts_per_million_expression_table),"adj.P.Val"]
```


#### __Combining Ca and Kp data into a single master table__
```{r combined master}
combined_ca_kp_expression_table_with_de_comp<-rbind.fill(ca_counts_per_million_expression_table_with_de_comp,kp_counts_per_million_expression_table_with_de_comp)

#Adding a column to include DEG classifications
combined_ca_kp_expression_table_with_de_comp$venn<-""
combined_ca_kp_expression_table_with_de_comp$venn[combined_ca_kp_expression_table_with_de_comp$locus_tag %in% signatures$CaKp_vs_Ca_RPMI_DEGS_UPREG] <- "RPMI_UP"
combined_ca_kp_expression_table_with_de_comp$venn[combined_ca_kp_expression_table_with_de_comp$locus_tag %in% signatures$CaKp_vs_Ca_RPMI_DEGS_DOWNREG] <- "RPMI_DOWN"
for (i in which(combined_ca_kp_expression_table_with_de_comp$locus_tag %in% signatures$CaKp_vs_Ca_M9_DEGS_UPREG)){
  combined_ca_kp_expression_table_with_de_comp$venn[i]<-ifelse(combined_ca_kp_expression_table_with_de_comp$venn[i] == "", "M9_UP", paste(combined_ca_kp_expression_table_with_de_comp$venn[i], "M9_UP", sep = ","))
}
for (i in which(combined_ca_kp_expression_table_with_de_comp$locus_tag %in% signatures$CaKp_vs_Ca_M9_DEGS_DOWNREG)){
  combined_ca_kp_expression_table_with_de_comp$venn[i]<-ifelse(combined_ca_kp_expression_table_with_de_comp$venn[i] == "", "M9_DOWN", paste(combined_ca_kp_expression_table_with_de_comp$venn[i], "M9_DOWN", sep = ","))
}


#Write table to file
write.xlsx(combined_ca_kp_expression_table_with_de_comp,file=file.path(OUTPUT_DIR,paste(deparse(substitute(combined_ca_kp_expression_table_with_de_comp)),".xlsx",sep="")),quote=FALSE,rowNames=FALSE,keepNA=TRUE,na.string="NA")
```


### __Single PCA for Candida and Klebsiella genes__

__Combining Ca top 1000 most variable genes and Kp top 1000 most variable genes for the single PCA plot__

> NA values present in Ca single or Kp single samples for Klebsiella and Candida genes respectively are replaced with zeroes

```{r combining top 1000}
ca_df<-as.data.frame(ca_top_lcpm)
kp_df<-as.data.frame(kp_top_lcpm)
combined_top_lcpm<-rbind.fill(ca_df,kp_df)
combined_top_lcpm[is.na(combined_top_lcpm)] <- 0
rownames(combined_top_lcpm)<-c(rownames(ca_top_lcpm),rownames(kp_top_lcpm))
```

```{r combined PCA generation, message=FALSE, eval=TRUE, include=TRUE}
pca_group<-gsub("/","",gsub(" mix","",gsub(" single","",paste(metadata$group_name_2,metadata$group_name_1))))
additional_col <- data.frame(culture = factor(metadata$group_name_2),media = factor(metadata$group_name_1), group = pca_group, stringsAsFactors =FALSE)
rownames(additional_col)<-gsub(" ", "", metadata$sample_name_in_report, fixed = TRUE)
additional_col<-additional_col[match(colnames(combined_top_lcpm),rownames(additional_col)),]
data_for_pca <- cbind(additional_col[colnames(combined_top_lcpm),],t(as.matrix(combined_top_lcpm)))
pca_object <- prcomp(data_for_pca[,-c(1:3)], center = TRUE, scale=TRUE)
summ<-summary(pca_object)
```

```{r combined pca plot, echo=FALSE,fig.height=7,fig.width=7}

par(mar=c(3,3,3,3))

cols<-c("#00AFBB","#00676F", "#E7B800","#002FE7", "#FC4E07","#BB0C00")
names(cols)<-levels(factor(pca_group))

pca_data<-additional_col
pca_data=cbind(pca_data,pca_object$x[,1],pca_object$x[,2])
colnames(pca_data)<-c(colnames(additional_col),"PC1","PC2")

# Compute group centroids for text labels
group_centroids <- pca_data %>%
  group_by(group) %>%
  dplyr::summarise(PC1 = mean(PC1), PC2 = mean(PC2))

# Calculate percentage variance explained
variance_explained <- (pca_object$sdev^2 / sum(pca_object$sdev^2)) * 100

# Create axis labels with variance percentages
pc1_label <- paste0("PC1 (", round(variance_explained[1], 2), "%)")
pc2_label <- paste0("PC2 (", round(variance_explained[2], 2), "%)")


ellipses <- pca_data %>%
  group_by(group) %>%
  group_map(~ {
    group_name <- .y$group[1]
    # Calculate the covariance matrix for each group
    cov_matrix <- cov_matrix <- cov(.[, c("PC1", "PC2")])
    cov_matrix[1, 1] <- cov_matrix[1, 1] +12
    cov_matrix[2, 2] <- cov_matrix[1, 1]+9
    # Calculate the mean for PC1 and PC2 (center of the ellipse)
    center <- colMeans(.[, c("PC1", "PC2")])  # Mean of PC1 and PC2
    center <- as.numeric(center)
    # Generate the ellipse data using the covariance matrix and the center
    ellipse_data <- car::ellipse(center = center, shape = cov_matrix, radius = 1.39, levels = 0.9,draw=FALSE)
    # Scale the ellipse by a factor (e.g., 1.2x)
    ellipse_data <- data.frame(ellipse_data)  # Convert to data frame
    ellipse_data$group <- group_name  # Add the group column
    return(ellipse_data)
  })
ellipses_df <- bind_rows(ellipses)
p<-ggplot(pca_data, aes(x = PC1, y = PC2, color = group)) +
  geom_point() +  # Keeps original points visible
  geom_path(data = ellipses_df, aes(x = x, y = y, group = group,color=group), size = 1) +
  labs(x = pc1_label, y = pc2_label, title = "PCA Plot") +
  scale_color_manual(values=cols) +
  theme_minimal() +
  theme(legend.position = "none",panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    geom_text_repel(data = group_centroids, aes(x = PC1, y = PC2, label = group), color = "black", fontface = "bold", box.padding = 3,segment.color = NA) +
  annotate(geom = 'segment', y = Inf, yend = Inf, color = 'black', x = -Inf, xend = Inf,size=1.1) +
  annotate(geom = 'segment', x = Inf, xend = Inf, color = 'black', y = -Inf, yend = Inf,size=1.1)
pdf(file.path(OUTPUT_DIR,"Ca_Kp_PCA.pdf"),height=6,width=6)
p
dev.off()
p

```


#### __PCA for RPMI samples__


```{r RPMI PCA generation, message=FALSE, eval=TRUE, include=TRUE}
data_for_pca_rpmi <- data_for_pca[which(data_for_pca$media=="RPMI"),] 
pca_object <- prcomp(data_for_pca_rpmi[,-c(1:3)], center = TRUE, scale=TRUE)
summ<-summary(pca_object)
```


```{r RPMI pca plot, echo=FALSE,fig.height=7,fig.width=7}

par(mar=c(3,3,3,3))
cols_rpmi<-cols[grep("RPMI",names(cols))]

pca_data<-additional_col[which(additional_col$media=="RPMI"),]
pca_data=cbind(pca_data,pca_object$x[,1],pca_object$x[,2])
colnames(pca_data)<-c(colnames(additional_col),"PC1","PC2")

# Compute group centroids for text labels
group_centroids <- pca_data %>%
  group_by(group) %>%
  dplyr::summarise(PC1 = mean(PC1), PC2 = mean(PC2))

# Calculate percentage variance explained
variance_explained <- (pca_object$sdev^2 / sum(pca_object$sdev^2)) * 100

# Create axis labels with variance percentages
pc1_label <- paste0("PC1 (", round(variance_explained[1], 2), "%)")
pc2_label <- paste0("PC2 (", round(variance_explained[2], 2), "%)")


ellipses <- pca_data %>%
  group_by(group) %>%
  group_map(~ {
    group_name <- .y$group[1]
    # Calculate the covariance matrix for each group
    cov_matrix <- cov_matrix <- cov(.[, c("PC1", "PC2")])
    cov_matrix[1, 1] <- cov_matrix[1, 1] +12
    cov_matrix[2, 2] <- cov_matrix[1, 1]+9
    # Calculate the mean for PC1 and PC2 (center of the ellipse)
    center <- colMeans(.[, c("PC1", "PC2")])  # Mean of PC1 and PC2
    center <- as.numeric(center)
    # Generate the ellipse data using the covariance matrix and the center
    ellipse_data <- car::ellipse(center = center, shape = cov_matrix, radius = 1.39, levels = 0.9,draw=FALSE)
    # Scale the ellipse by a factor (e.g., 1.2x)
    ellipse_data <- data.frame(ellipse_data)  # Convert to data frame
    ellipse_data$group <- group_name  # Add the group column
    return(ellipse_data)
  })
ellipses_df <- bind_rows(ellipses)
p<-ggplot(pca_data, aes(x = PC1, y = PC2, color = group)) +
  geom_point() +  # Keeps original points visible
  geom_path(data = ellipses_df, aes(x = x, y = y, group = group,color=group), size = 1) +
  labs(x = pc1_label, y = pc2_label, title = "PCA Plot") +
  scale_color_manual(values=cols_rpmi) +
  theme_minimal() +
  theme(legend.position = "none",panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    geom_text_repel(data = group_centroids, aes(x = PC1, y = PC2, label = group), color = "black", fontface = "bold", box.padding = 3,segment.color = NA) +
  annotate(geom = 'segment', y = Inf, yend = Inf, color = 'black', x = -Inf, xend = Inf,size=1.1) +
  annotate(geom = 'segment', x = Inf, xend = Inf, color = 'black', y = -Inf, yend = Inf,size=1.1)
pdf(file.path(OUTPUT_DIR,"Ca_Kp_PCA_RPMI.pdf"),height=6,width=6)
p
dev.off()
p

```

#### __PCA for M9 samples__

```{r M9 PCA generation, message=FALSE, eval=TRUE, include=TRUE}
data_for_pca_m9 <- data_for_pca[which(data_for_pca$media=="M9"),] 
pca_object <- prcomp(data_for_pca_m9[,-c(1:3)], center = TRUE, scale=TRUE)
summ<-summary(pca_object)
```

```{r M9 pca plot, echo=FALSE,fig.height=7,fig.width=7}

par(mar=c(3,3,3,3))
cols_rpmi<-cols[grep("M9",names(cols))]

pca_data<-additional_col[which(additional_col$media=="M9"),]
pca_data=cbind(pca_data,pca_object$x[,1],pca_object$x[,2])
colnames(pca_data)<-c(colnames(additional_col),"PC1","PC2")

# Compute group centroids for text labels
group_centroids <- pca_data %>%
  group_by(group) %>%
  dplyr::summarise(PC1 = mean(PC1), PC2 = mean(PC2))

# Calculate percentage variance explained
variance_explained <- (pca_object$sdev^2 / sum(pca_object$sdev^2)) * 100

# Create axis labels with variance percentages
pc1_label <- paste0("PC1 (", round(variance_explained[1], 2), "%)")
pc2_label <- paste0("PC2 (", round(variance_explained[2], 2), "%)")


ellipses <- pca_data %>%
  group_by(group) %>%
  group_map(~ {
    group_name <- .y$group[1]
    # Calculate the covariance matrix for each group
    cov_matrix <- cov_matrix <- cov(.[, c("PC1", "PC2")])
    cov_matrix[1, 1] <- cov_matrix[1, 1] +12
    cov_matrix[2, 2] <- cov_matrix[1, 1]+9
    # Calculate the mean for PC1 and PC2 (center of the ellipse)
    center <- colMeans(.[, c("PC1", "PC2")])  # Mean of PC1 and PC2
    center <- as.numeric(center)
    # Generate the ellipse data using the covariance matrix and the center
    ellipse_data <- car::ellipse(center = center, shape = cov_matrix, radius = 1.39, levels = 0.9,draw=FALSE)
    # Scale the ellipse by a factor (e.g., 1.2x)
    ellipse_data <- data.frame(ellipse_data)  # Convert to data frame
    ellipse_data$group <- group_name  # Add the group column
    return(ellipse_data)
  })
ellipses_df <- bind_rows(ellipses)
p<-ggplot(pca_data, aes(x = PC1, y = PC2, color = group)) +
  geom_point() +  # Keeps original points visible
  geom_path(data = ellipses_df, aes(x = x, y = y, group = group,color=group), size = 1) +
  labs(x = pc1_label, y = pc2_label, title = "PCA Plot") +
  scale_color_manual(values=cols_rpmi) +
  theme_minimal() +
  theme(legend.position = "none",panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    geom_text_repel(data = group_centroids, aes(x = PC1, y = PC2, label = group), color = "black", fontface = "bold", box.padding = 3,segment.color = NA) +
  annotate(geom = 'segment', y = Inf, yend = Inf, color = 'black', x = -Inf, xend = Inf,size=1.1) +
  annotate(geom = 'segment', x = Inf, xend = Inf, color = 'black', y = -Inf, yend = Inf,size=1.1)
pdf(file.path(OUTPUT_DIR,"Ca_Kp_PCA_M9.pdf"),height=6,width=6)
p
dev.off()
p

```

### __Replot PCA for Candida genes__


```{r Ca combined PCA generation, message=FALSE, eval=TRUE, include=TRUE}
additional_col_ca <- additional_col[grep("C",data_for_pca$group),]
data_for_pca_ca <- cbind(additional_col[colnames(ca_top_lcpm),],t(as.matrix(ca_top_lcpm)))
pca_object <- prcomp(data_for_pca_ca[,-c(1:3)], center = TRUE, scale=TRUE)
summ<-summary(pca_object)
```

```{r Ca combined pca plot, echo=FALSE,fig.height=7,fig.width=7}

par(mar=c(3,3,3,3))

pca_data<-additional_col_ca
pca_data=cbind(pca_data,pca_object$x[,1],pca_object$x[,2])
colnames(pca_data)<-c(colnames(additional_col_ca),"PC1","PC2")

# Compute group centroids for text labels
group_centroids <- pca_data %>%
  group_by(group) %>%
  dplyr::summarise(PC1 = mean(PC1), PC2 = mean(PC2))

# Calculate percentage variance explained
variance_explained <- (pca_object$sdev^2 / sum(pca_object$sdev^2)) * 100

# Create axis labels with variance percentages
pc1_label <- paste0("PC1 (", round(variance_explained[1], 2), "%)")
pc2_label <- paste0("PC2 (", round(variance_explained[2], 2), "%)")


ellipses <- pca_data %>%
  group_by(group) %>%
  group_map(~ {
    group_name <- .y$group[1]
    # Calculate the covariance matrix for each group
    cov_matrix <- cov_matrix <- cov(.[, c("PC1", "PC2")])
    cov_matrix[1, 1] <- cov_matrix[1, 1] +12
    cov_matrix[2, 2] <- cov_matrix[1, 1]+9
    # Calculate the mean for PC1 and PC2 (center of the ellipse)
    center <- colMeans(.[, c("PC1", "PC2")])  # Mean of PC1 and PC2
    center <- as.numeric(center)
    # Generate the ellipse data using the covariance matrix and the center
    ellipse_data <- car::ellipse(center = center, shape = cov_matrix, radius = 1.39, levels = 0.9,draw=FALSE)
    # Scale the ellipse by a factor (e.g., 1.2x)
    ellipse_data <- data.frame(ellipse_data)  # Convert to data frame
    ellipse_data$group <- group_name  # Add the group column
    return(ellipse_data)
  })
ellipses_df <- bind_rows(ellipses)
p<-ggplot(pca_data, aes(x = PC1, y = PC2, color = group)) +
  geom_point() +  # Keeps original points visible
  geom_path(data = ellipses_df, aes(x = x, y = y, group = group,color=group), size = 1) +
  labs(x = pc1_label, y = pc2_label, title = "PCA Plot") +
  scale_color_manual(values=cols) +
  theme_minimal() +
  theme(legend.position = "none",panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    geom_text_repel(data = group_centroids, aes(x = PC1, y = PC2, label = group), color = "black", fontface = "bold", box.padding = 3,segment.color = NA) +
  annotate(geom = 'segment', y = Inf, yend = Inf, color = 'black', x = -Inf, xend = Inf,size=1.1) +
  annotate(geom = 'segment', x = Inf, xend = Inf, color = 'black', y = -Inf, yend = Inf,size=1.1)
pdf(file.path(OUTPUT_DIR,"Ca_PCA.pdf"),height=6,width=6)
p
dev.off()
p

```


#### __PCA for Ca genes RPMI media samples__


```{r Ca RPMI PCA generation, message=FALSE, eval=TRUE, include=TRUE}
data_for_pca_ca_rpmi <- data_for_pca_ca[which(data_for_pca_ca$media=="RPMI"),] 
pca_object <- prcomp(data_for_pca_ca_rpmi[,-c(1:3)], center = TRUE, scale=TRUE)
summ<-summary(pca_object)
```


```{r Ca RPMI pca plot, echo=FALSE,fig.height=7,fig.width=7}

par(mar=c(3,3,3,3))
cols_rpmi<-cols[grep("RPMI",names(cols))]

pca_data<-additional_col_ca[which(additional_col_ca$media=="RPMI"),]
pca_data=cbind(pca_data,pca_object$x[,1],pca_object$x[,2])
colnames(pca_data)<-c(colnames(additional_col_ca),"PC1","PC2")

# Compute group centroids for text labels
group_centroids <- pca_data %>%
  group_by(group) %>%
  dplyr::summarise(PC1 = mean(PC1), PC2 = mean(PC2))

# Calculate percentage variance explained
variance_explained <- (pca_object$sdev^2 / sum(pca_object$sdev^2)) * 100

# Create axis labels with variance percentages
pc1_label <- paste0("PC1 (", round(variance_explained[1], 2), "%)")
pc2_label <- paste0("PC2 (", round(variance_explained[2], 2), "%)")


ellipses <- pca_data %>%
  group_by(group) %>%
  group_map(~ {
    group_name <- .y$group[1]
    # Calculate the covariance matrix for each group
    cov_matrix <- cov_matrix <- cov(.[, c("PC1", "PC2")])
    # Reduce tilt
    cov_matrix[1, 2] <- cov_matrix[1, 2] * 0.7
    cov_matrix[2, 1] <- cov_matrix[1, 2]
    # Make ellipse flatter
    cov_matrix[1, 1] <- cov_matrix[1, 1] * 1.3   # widen horizontal (optional)
    # Calculate the mean for PC1 and PC2 (center of the ellipse)
    center <- colMeans(.[, c("PC1", "PC2")])  # Mean of PC1 and PC2
    center <- as.numeric(center)
    # Generate the ellipse data using the covariance matrix and the center
    ellipse_data <- car::ellipse(center = center, shape = cov_matrix, radius = 1.39, levels = 0.9,draw=FALSE)
    # Scale the ellipse by a factor (e.g., 1.2x)
    ellipse_data <- data.frame(ellipse_data)  # Convert to data frame
    ellipse_data$group <- group_name  # Add the group column
    return(ellipse_data)
  })
ellipses_df <- bind_rows(ellipses)
p<-ggplot(pca_data, aes(x = PC1, y = PC2, color = group)) +
  geom_point() +  # Keeps original points visible
  geom_path(data = ellipses_df, aes(x = x, y = y, group = group,color=group), size = 1) +
  labs(x = pc1_label, y = pc2_label, title = "PCA Plot") +
  scale_color_manual(values=cols_rpmi) +
  theme_minimal() +
  theme(legend.position = "none",panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    geom_text_repel(data = group_centroids, aes(x = PC1, y = PC2, label = group), color = "black", fontface = "bold", box.padding = 3,segment.color = NA) +
  annotate(geom = 'segment', y = Inf, yend = Inf, color = 'black', x = -Inf, xend = Inf,size=1.1) +
  annotate(geom = 'segment', x = Inf, xend = Inf, color = 'black', y = -Inf, yend = Inf,size=1.1)
pdf(file.path(OUTPUT_DIR,"Ca_PCA_RPMI.pdf"),height=6,width=6)
p
dev.off()
p

```

#### __PCA for Ca genes M9 media samples__

```{r Ca M9 PCA generation, message=FALSE, eval=TRUE, include=TRUE}
data_for_pca_ca_m9 <- data_for_pca_ca[which(data_for_pca_ca$media=="M9"),] 
pca_object <- prcomp(data_for_pca_ca_m9[,-c(1:3)], center = TRUE, scale=TRUE)
summ<-summary(pca_object)
```

```{r Ca M9 pca plot, echo=FALSE,fig.height=7,fig.width=7}

par(mar=c(3,3,3,3))
cols_rpmi<-cols[grep("M9",names(cols))]

pca_data<-additional_col_ca[which(additional_col_ca$media=="M9"),]
pca_data=cbind(pca_data,pca_object$x[,1],pca_object$x[,2])
colnames(pca_data)<-c(colnames(additional_col_ca),"PC1","PC2")

# Compute group centroids for text labels
group_centroids <- pca_data %>%
  group_by(group) %>%
  dplyr::summarise(PC1 = mean(PC1), PC2 = mean(PC2))

# Calculate percentage variance explained
variance_explained <- (pca_object$sdev^2 / sum(pca_object$sdev^2)) * 100

# Create axis labels with variance percentages
pc1_label <- paste0("PC1 (", round(variance_explained[1], 2), "%)")
pc2_label <- paste0("PC2 (", round(variance_explained[2], 2), "%)")


ellipses <- pca_data %>%
  group_by(group) %>%
  group_map(~ {
    group_name <- .y$group[1]
    # Calculate the covariance matrix for each group
    cov_matrix <- cov_matrix <- cov(.[, c("PC1", "PC2")])
    # Reduce tilt
    cov_matrix[1, 2] <- cov_matrix[1, 2] * 0.3
    cov_matrix[2, 1] <- cov_matrix[1, 2]
    # Increase vertical spread
    cov_matrix[2, 2] <- cov_matrix[2, 2] * 2
    # Calculate the mean for PC1 and PC2 (center of the ellipse)
    center <- colMeans(.[, c("PC1", "PC2")])  # Mean of PC1 and PC2
    center <- as.numeric(center)
    # Generate the ellipse data using the covariance matrix and the center
    ellipse_data <- car::ellipse(center = center, shape = cov_matrix, radius = 4, levels = 0.9,draw=FALSE)
    # Scale the ellipse by a factor (e.g., 1.2x)
    ellipse_data <- data.frame(ellipse_data)  # Convert to data frame
    ellipse_data$group <- group_name  # Add the group column
    return(ellipse_data)
  })
ellipses_df <- bind_rows(ellipses)
p<-ggplot(pca_data, aes(x = PC1, y = PC2, color = group)) +
  geom_point() +  # Keeps original points visible
  geom_path(data = ellipses_df, aes(x = x, y = y, group = group,color=group), size = 1) +
  labs(x = pc1_label, y = pc2_label, title = "PCA Plot") +
  scale_color_manual(values=cols_rpmi) +
  theme_minimal() +
  theme(legend.position = "none",panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    geom_text_repel(data = group_centroids, aes(x = PC1, y = PC2, label = group), color = "black", fontface = "bold", box.padding = 3,segment.color = NA) +
  annotate(geom = 'segment', y = Inf, yend = Inf, color = 'black', x = -Inf, xend = Inf,size=1.1) +
  annotate(geom = 'segment', x = Inf, xend = Inf, color = 'black', y = -Inf, yend = Inf,size=1.1)
pdf(file.path(OUTPUT_DIR,"Ca_PCA_M9.pdf"),height=6,width=6)
p
dev.off()
p

```

### __Replot PCA for Klebsiella genes__


```{r Kp combined PCA generation, message=FALSE, eval=TRUE, include=TRUE}
additional_col_kp <- additional_col[grep("K",data_for_pca$group),]
data_for_pca_kp <- cbind(additional_col[colnames(kp_top_lcpm),],t(as.matrix(kp_top_lcpm)))
pca_object <- prcomp(data_for_pca_kp[,-c(1:3)], center = TRUE, scale=TRUE)
summ<-summary(pca_object)
```

```{r Kp combined pca plot, echo=FALSE,fig.height=7,fig.width=7}

par(mar=c(3,3,3,3))

pca_data<-additional_col_kp
pca_data=cbind(pca_data,pca_object$x[,1],pca_object$x[,2])
colnames(pca_data)<-c(colnames(additional_col_kp),"PC1","PC2")

# Compute group centroids for text labels
group_centroids <- pca_data %>%
  group_by(group) %>%
  dplyr::summarise(PC1 = mean(PC1), PC2 = mean(PC2))

# Calculate percentage variance explained
variance_explained <- (pca_object$sdev^2 / sum(pca_object$sdev^2)) * 100

# Create axis labels with variance percentages
pc1_label <- paste0("PC1 (", round(variance_explained[1], 2), "%)")
pc2_label <- paste0("PC2 (", round(variance_explained[2], 2), "%)")


ellipses <- pca_data %>%
  group_by(group) %>%
  group_map(~ {
    group_name <- .y$group[1]
    # Calculate the covariance matrix for each group
    cov_matrix <- cov_matrix <- cov(.[, c("PC1", "PC2")])
    cov_matrix[1, 1] <- cov_matrix[1, 1] +12
    cov_matrix[2, 2] <- cov_matrix[1, 1]+9
    # Calculate the mean for PC1 and PC2 (center of the ellipse)
    center <- colMeans(.[, c("PC1", "PC2")])  # Mean of PC1 and PC2
    center <- as.numeric(center)
    # Generate the ellipse data using the covariance matrix and the center
    ellipse_data <- car::ellipse(center = center, shape = cov_matrix, radius = 1.39, levels = 0.9,draw=FALSE)
    # Scale the ellipse by a factor (e.g., 1.2x)
    ellipse_data <- data.frame(ellipse_data)  # Convert to data frame
    ellipse_data$group <- group_name  # Add the group column
    return(ellipse_data)
  })
ellipses_df <- bind_rows(ellipses)
p<-ggplot(pca_data, aes(x = PC1, y = PC2, color = group)) +
  geom_point() +  # Keeps original points visible
  geom_path(data = ellipses_df, aes(x = x, y = y, group = group,color=group), size = 1) +
  labs(x = pc1_label, y = pc2_label, title = "PCA Plot") +
  scale_color_manual(values=cols) +
  theme_minimal() +
  theme(legend.position = "none",panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    geom_text_repel(data = group_centroids, aes(x = PC1, y = PC2, label = group), color = "black", fontface = "bold", box.padding = 1.5,segment.color = NA) +
  annotate(geom = 'segment', y = Inf, yend = Inf, color = 'black', x = -Inf, xend = Inf,size=1.1) +
  annotate(geom = 'segment', x = Inf, xend = Inf, color = 'black', y = -Inf, yend = Inf,size=1.1)
pdf(file.path(OUTPUT_DIR,"Kp_PCA.pdf"),height=6,width=6)
p
dev.off()
p

```


#### __PCA for Kp genes RPMI media samples__


```{r Kp RPMI PCA generation, message=FALSE, eval=TRUE, include=TRUE}
data_for_pca_kp_rpmi <- data_for_pca_kp[which(data_for_pca_kp$media=="RPMI"),] 
pca_object <- prcomp(data_for_pca_kp_rpmi[,-c(1:3)], center = TRUE, scale=TRUE)
summ<-summary(pca_object)
```


```{r Kp RPMI pca plot, echo=FALSE,fig.height=7,fig.width=7}

par(mar=c(3,3,3,3))
cols_rpmi<-cols[grep("RPMI",names(cols))]

pca_data<-additional_col_kp[which(additional_col_kp$media=="RPMI"),]
pca_data=cbind(pca_data,pca_object$x[,1],pca_object$x[,2])
colnames(pca_data)<-c(colnames(additional_col_kp),"PC1","PC2")

# Compute group centroids for text labels
group_centroids <- pca_data %>%
  group_by(group) %>%
  dplyr::summarise(PC1 = mean(PC1), PC2 = mean(PC2))

# Calculate percentage variance explained
variance_explained <- (pca_object$sdev^2 / sum(pca_object$sdev^2)) * 100

# Create axis labels with variance percentages
pc1_label <- paste0("PC1 (", round(variance_explained[1], 2), "%)")
pc2_label <- paste0("PC2 (", round(variance_explained[2], 2), "%)")


ellipses <- pca_data %>%
  group_by(group) %>%
  group_map(~ {
    group_name <- .y$group[1]
    # Calculate the covariance matrix for each group
    cov_matrix <- cov_matrix <- cov(.[, c("PC1", "PC2")])
    # Reduce tilt
    cov_matrix[1, 2] <- cov_matrix[1, 2] * 0.7
    cov_matrix[2, 1] <- cov_matrix[1, 2]
    # Make ellipse flatter
    cov_matrix[1, 1] <- cov_matrix[1, 1] * 1.3   # widen horizontal (optional)
    # Calculate the mean for PC1 and PC2 (center of the ellipse)
    center <- colMeans(.[, c("PC1", "PC2")])  # Mean of PC1 and PC2
    center <- as.numeric(center)
    # Generate the ellipse data using the covariance matrix and the center
    ellipse_data <- car::ellipse(center = center, shape = cov_matrix, radius = 2, levels = 0.9,draw=FALSE)
    # Scale the ellipse by a factor (e.g., 1.2x)
    ellipse_data <- data.frame(ellipse_data)  # Convert to data frame
    ellipse_data$group <- group_name  # Add the group column
    return(ellipse_data)
  })
ellipses_df <- bind_rows(ellipses)
p<-ggplot(pca_data, aes(x = PC1, y = PC2, color = group)) +
  geom_point() +  # Keeps original points visible
  geom_path(data = ellipses_df, aes(x = x, y = y, group = group,color=group), size = 1) +
  labs(x = pc1_label, y = pc2_label, title = "PCA Plot") +
  scale_color_manual(values=cols_rpmi) +
  theme_minimal() +
  theme(legend.position = "none",panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    geom_text_repel(data = group_centroids, aes(x = PC1, y = PC2, label = group), color = "black", fontface = "bold", box.padding = 3,segment.color = NA) +
  annotate(geom = 'segment', y = Inf, yend = Inf, color = 'black', x = -Inf, xend = Inf,size=1.1) +
  annotate(geom = 'segment', x = Inf, xend = Inf, color = 'black', y = -Inf, yend = Inf,size=1.1)
pdf(file.path(OUTPUT_DIR,"Kp_PCA_RPMI.pdf"),height=6,width=6)
p
dev.off()
p

```

#### __PCA for Kp genes M9 media samples__

```{r Kp M9 PCA generation, message=FALSE, eval=TRUE, include=TRUE}
data_for_pca_kp_m9 <- data_for_pca_kp[which(data_for_pca_kp$media=="M9"),] 
pca_object <- prcomp(data_for_pca_kp_m9[,-c(1:3)], center = TRUE, scale=TRUE)
summ<-summary(pca_object)
```

```{r Kp M9 pca plot, echo=FALSE,fig.height=7,fig.width=7}

par(mar=c(3,3,3,3))
cols_rpmi<-cols[grep("M9",names(cols))]

pca_data<-additional_col_kp[which(additional_col_kp$media=="M9"),]
pca_data=cbind(pca_data,pca_object$x[,1],pca_object$x[,2])
colnames(pca_data)<-c(colnames(additional_col_kp),"PC1","PC2")

# Compute group centroids for text labels
group_centroids <- pca_data %>%
  group_by(group) %>%
  dplyr::summarise(PC1 = mean(PC1), PC2 = mean(PC2))

# Calculate percentage variance explained
variance_explained <- (pca_object$sdev^2 / sum(pca_object$sdev^2)) * 100

# Create axis labels with variance percentages
pc1_label <- paste0("PC1 (", round(variance_explained[1], 2), "%)")
pc2_label <- paste0("PC2 (", round(variance_explained[2], 2), "%)")


ellipses <- pca_data %>%
  group_by(group) %>%
  group_map(~ {
    group_name <- .y$group[1]
    # Calculate the covariance matrix for each group
    cov_matrix <- cov_matrix <- cov(.[, c("PC1", "PC2")])
    # Reduce tilt
    cov_matrix[1, 2] <- cov_matrix[1, 2] * 0.7
    cov_matrix[2, 1] <- cov_matrix[1, 2]
    # Make ellipse flatter
    cov_matrix[2, 2] <- cov_matrix[2, 2] * 0.7   # shrink vertical
    cov_matrix[1, 1] <- cov_matrix[1, 1] * 1.3   # widen horizontal (optional)
    # Calculate the mean for PC1 and PC2 (center of the ellipse)
    center <- colMeans(.[, c("PC1", "PC2")])  # Mean of PC1 and PC2
    center <- as.numeric(center)
    # Generate the ellipse data using the covariance matrix and the center
    ellipse_data <- car::ellipse(center = center, shape = cov_matrix, radius = 1.8, levels = 0.9,draw=FALSE)
    # Scale the ellipse by a factor (e.g., 1.2x)
    ellipse_data <- data.frame(ellipse_data)  # Convert to data frame
    ellipse_data$group <- group_name  # Add the group column
    return(ellipse_data)
  })
ellipses_df <- bind_rows(ellipses)
p<-ggplot(pca_data, aes(x = PC1, y = PC2, color = group)) +
  geom_point() +  # Keeps original points visible
  geom_path(data = ellipses_df, aes(x = x, y = y, group = group,color=group), size = 1) +
  labs(x = pc1_label, y = pc2_label, title = "PCA Plot") +
  scale_color_manual(values=cols_rpmi) +
  theme_minimal() +
  theme(legend.position = "none",panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
    geom_text_repel(data = group_centroids, aes(x = PC1, y = PC2, label = group), color = "black", fontface = "bold", box.padding = 3,segment.color = NA) +
  annotate(geom = 'segment', y = Inf, yend = Inf, color = 'black', x = -Inf, xend = Inf,size=1.1) +
  annotate(geom = 'segment', x = Inf, xend = Inf, color = 'black', y = -Inf, yend = Inf,size=1.1)
pdf(file.path(OUTPUT_DIR,"Kp_PCA_M9.pdf"),height=6,width=6)
p
dev.off()
p

```